---
title: "Individual differences in habituation rate and behavioral volatility predict dishabituation in adults, preschoolers and infants"
author: "Anonymous CogSci submission"
date: "`r Sys.Date()`"
output:
  pdf_document: 
    toc: true
  html_document: default
subtitle: "Supplementary Information" 

---

```{r include = FALSE}
knitr::opts_chunk$set(echo=FALSE, message = FALSE, warning = FALSE)
```


```{r}
library(tidyverse)
library(here)
library(corrplot)
library(kableExtra)
library(ggthemes)

cormatrix_adult <- readRDS(here('cached_results/cormatrix_adult.Rds'))
cormatrix_preschooler <- readRDS(here('cached_results/cormatrix_preschooler.Rds'))
cormatrix_infant <- readRDS(here('cached_results/cormatrix_infant.Rds'))

pmatrix_adult <- readRDS(here("cached_results/pmatrix_adult.Rds"))
pmatrix_preschooler <- readRDS(here("cached_results/pmatrix_preschooler.Rds"))
pmatrix_infant <- readRDS(here("cached_results/pmatrix_infant.Rds"))

model_comp_resid_based_adult <- readRDS(here("cached_results/model_comp_resid_based_adult.Rds"))
model_comp_resid_based_preschooler <- readRDS(here("cached_results/model_comp_resid_based_preschooler.Rds"))
model_comp_resid_based_infant <- readRDS(here("cached_results/model_comp_resid_based_infant.Rds"))

model_comp_logdiff_adult <- readRDS(here("cached_results/model_comp_logdiff_adult.Rds"))
model_comp_logdiff_preschooler <- readRDS(here("cached_results/model_comp_logdiff_preschooler.Rds"))
model_comp_logdiff_infant <- readRDS(here("cached_results/model_comp_logdiff_infant.Rds"))
```



# Simulations 

## Generating dataset 

```{r }
library(ggplot2)
library(dplyr)
library(tidyr)

set.seed(42)

#### 1. Generate Simulated Data ####
N <- 200                # Number of participants
n_hab <- 5              # Number of habituation trials
true_a <- 1.5           # Population intercept for habituation
true_b <- -0.3          # Population slope for habituation
sigma_baseline <- 0.5   # SD of random intercepts
sigma_slope <- 0.2      # SD of random slopes
sigma_obs <- 0.2        # Observation noise
alpha_dishab <- 3.0     # Dishabituation intercept
true_dishab <- 2.0      # "True" dishabituation effect
sigma_dishab <- 0.3     # Noise on dishabituation

# Generate per-participant random effects
a_i <- rnorm(N, 0, sigma_baseline)  # Random intercepts
b_i <- rnorm(N, 0, sigma_slope)     # Random slopes

# Create empty dataframe
sim_data <- data.frame(
  id = rep(1:N, each = n_hab + 1),
  trial = rep(1:(n_hab + 1), times = N),
  t_ij = NA_real_,
  baseline = rep(a_i, each = n_hab + 1)
)

# Generate habituation trials
for(i in 1:N){
  for(j in 1:n_hab){
    row_index <- (i-1)*(n_hab+1) + j
    sim_data$t_ij[row_index] <- exp(
      (true_a + a_i[i]) + (true_b + b_i[i])*j
    ) + rnorm(1, 0, sigma_obs)
  }
}

# Generate dishabituation trial (novel stimulus)
for(i in 1:N){
  row_index <- (i-1)*(n_hab+1) + (n_hab+1)
  sim_data$t_ij[row_index] <- 
    (alpha_dishab + a_i[i] + true_dishab) + rnorm(1, 0, sigma_dishab)
}
```


Here we simulated a dataset to model habituation and dishabituation effects in a looking time paradigm. Each of the 200 participants undergoes five habituation trials followed by one dishabituation trial. The habituation process is modeled using an exponential decay function, where looking time decreases across trials according to a population-level intercept and slope, with individual variability introduced through random intercepts and slopes. Observation noise is added to simulate natural variability in responses. The dishabituation trial, representing a response to a novel stimulus, is generated using a separate intercept with an additional dishabituation effect and individual-level noise. This setup captures both the within-subject decline in looking time during habituation and the rebound effect when a novel stimulus is introduced.

```{r fig.pos="t!", fig.width = 4, fig.height = 4}
sim_data %>% 
  ggplot(aes(x = trial, y = t_ij)) + 
  geom_point(position = position_jitter(width = .2), alpha = .2) + 
  stat_summary(fun.data = "mean_cl_boot", color = "red") + 
  theme_few()
```

```{r}
# Extract last habituation trial and dishabituation trial
classical_measures <- sim_data %>%
  filter(trial %in% c(n_hab, n_hab + 1)) %>%
  pivot_wider(
    id_cols = id,
    names_from = trial,
    values_from = t_ij,
    names_prefix = "trial_"
  ) %>%
  mutate(
    raw_dishabituation = trial_6,   # Raw looking time on dishabituation trial
    classical_val = log(trial_6) - log(trial_5)  # Log difference score
  ) %>%
  left_join(
    sim_data %>%
      distinct(id, .keep_all = TRUE) %>%
      select(id, baseline),
    by = "id"
  )

# Residual-Based Measure
# Fit an intercept-only model for dishabituation
dishab_data <- sim_data %>% filter(trial == n_hab + 1)
lm_dishab <- lm(t_ij ~ 1, data = dishab_data)
dishab_data$resid_dishab <- resid(lm_dishab)

#### 2. Compute Classical and Residual-Based Measures ####

# Fit a linear model for habituation
hab_data <- sim_data %>% filter(trial <= n_hab)
lm_hab <- lm(t_ij ~ trial, data = hab_data)

# Get residuals for the last habituation trial only
resid_last_hab <- hab_data %>%
  mutate(resid_hab = resid(lm_hab)) %>%
  filter(trial == n_hab) %>%
  select(id, resid_hab)

# Compute residual-based measure
residual_measures <- dishab_data %>%
  left_join(resid_last_hab, by = "id") %>%
  mutate(V_i = resid_dishab - resid_hab) %>%
  select(id, V_i)

# Merge all measures
final_data <- classical_measures %>%
  left_join(residual_measures, by = "id")

cor_raw_dishab <- cor(final_data$raw_dishabituation, final_data$baseline, use = "complete.obs")
cor_residual_based <- cor(final_data$V_i, final_data$baseline, use = "complete.obs")
```

## Comparison with baseline 

Comparison of raw dishabituation and residual-based measures against baseline looking time. The left panel displays the relationship between raw dishabituation looking time (i.e., looking time on the first post-habituation trial) and baseline looking time (*r* = `r round(cor_raw_dishab, 2)`), while the right panel shows the relationship between the residual-based measure ($V_i$; *r* = `r round(cor_residual_based, 2)`), which accounts for individual differences in habituation trends, and baseline looking time. Both measures exhibit variability in their associations with baseline looking, highlighting the impact of measurement choice on capturing dishabituation effects. Shaded areas represent 95% confidence intervals from linear model fits.

The plot shows that residual-based approach significantly reduced the correlation between dishabituation and the baseline looking time. 

```{r fig.pos="h!", fig.width = 8, fig.height = 4}
final_data %>% 
  pivot_longer(cols = c("raw_dishabituation", "V_i"), 
               names_to = "type", values_to = "value") %>% 
  mutate(
    type_clean = if_else(type == "raw_dishabituation", "Raw Dishabituation", "Residual-based Measure")
  )%>% 
  ggplot(aes(x = baseline, y = value)) + 
  geom_point() +
  geom_smooth(method = "lm") + 
  facet_wrap(~type_clean, scales = "free") + 
  theme_few() + 
  ylab("")


```





# Model comparisons 

## Residual-based models 

```{r fi}
model_comp_logdiff_adult %>% select(-group) %>% 
  kable("latex", booktabs = TRUE, align = "c", 
        caption = "Adults") %>%
  kable_styling(full_width = TRUE, position = "left") %>%
  kable_styling(latex_options = "HOLD_position")
```


```{r}
model_comp_logdiff_preschooler %>% select(-group) %>% 
  kable("latex", booktabs = TRUE, align = "c", 
        caption = "Preschoolers") %>%
  kable_styling(full_width = TRUE, position = "left") %>%
  kable_styling(latex_options = "HOLD_position")
```


```{r}
model_comp_logdiff_infant %>% select(-group) %>% 
  kable("latex", booktabs = TRUE, align = "c", 
        caption = "Infants") %>%
  kable_styling(full_width = TRUE, position = "left") %>%
  kable_styling(latex_options = "HOLD_position")
```


## Robustness check models

```{r}
model_comp_logdiff_adult %>% select(-group) %>% 
  kable("latex", booktabs = TRUE, align = "c", 
        caption = "Adults") %>%
  kable_styling(full_width = TRUE, position = "left") %>%
  kable_styling(latex_options = "HOLD_position")
```


```{r}
model_comp_logdiff_preschooler %>% select(-group) %>% 
  kable("latex", booktabs = TRUE, align = "c", 
        caption = "Preschoolers") %>%
  kable_styling(full_width = TRUE, position = "left") %>%
  kable_styling(latex_options = "HOLD_position")

```


```{r}
model_comp_logdiff_infant %>% select(-group) %>% 
  kable("latex", booktabs = TRUE, align = "c", 
        caption = "Infants") %>%
  kable_styling(full_width = TRUE, position = "left") %>%
  kable_styling(latex_options = "HOLD_position")
```

# Correlations between measures and predictors 

Below are the correlation matrices for all predictors and the two operationalizations of dishabituation. Specifically, `resid_diff` represents the residual-based dishabituation, and `log_diff` is the difference score between the log-transformed looking time at the dishabituation trial and the last habituation trial. The numbers in the matrices are Pearson’s correlation coefficients, and blank cells indicate nonsignificant correlations.



```{r fig.pos="H!", fig.width = 3.5, fig.height = 3.5, fig.cap="Correlational matrix for adults."}
corrplot(cormatrix_adult, 
         p.mat = pmatrix_adult, 
         method = "number",sig.level = 0.05, insig = "blank", title = "Adults", 
         mar = c(0, 0, 1, 0)) 
```


```{r fig.pos="H!", fig.width = 3.5, fig.height = 3.5, fig.cap="Correlational matrix for preschoolers. "}
corrplot(cormatrix_preschooler, 
         p.mat = pmatrix_preschooler, 
         method = "number",sig.level = 0.05, insig = "blank", title = "Preschoolers", 
         mar = c(0, 0, 1, 0)) 
```



```{r fig.pos="H!", fig.width = 3.5, fig.height = 3.5, fig.cap="Correlational matrix for infants."}
corrplot(cormatrix_infant, 
         p.mat = pmatrix_infant, 
         method = "number",sig.level = 0.05, insig = "blank", title = "Infants", 
         mar = c(0, 0, 1, 0)) 
```









