---
title: "Individual differences in habituation rate and behavioral volatility predict dishabituation in adults, preschoolers and infants"
author: "Anonymous CogSci submission"
date: "`r Sys.Date()`"
output:
  pdf_document: 
    toc: true
  html_document: default
subtitle: "Supplementary Information" 

---

```{r include = FALSE}
knitr::opts_chunk$set(echo=FALSE, message = FALSE)
```


```{r}
library(tidyverse)
library(here)
library(corrplot)
library(kableExtra)

cormatrix_adult <- readRDS(here('cached_results/cormatrix_adult.Rds'))
cormatrix_preschooler <- readRDS(here('cached_results/cormatrix_preschooler.Rds'))
cormatrix_infant <- readRDS(here('cached_results/cormatrix_infant.Rds'))

pmatrix_adult <- readRDS(here("cached_results/pmatrix_adult.Rds"))
pmatrix_preschooler <- readRDS(here("cached_results/pmatrix_preschooler.Rds"))
pmatrix_infant <- readRDS(here("cached_results/pmatrix_infant.Rds"))

model_comp_resid_based_adult <- readRDS(here("cached_results/model_comp_resid_based_adult.Rds"))
model_comp_resid_based_preschooler <- readRDS(here("cached_results/model_comp_resid_based_preschooler.Rds"))
model_comp_resid_based_infant <- readRDS(here("cached_results/model_comp_resid_based_infant.Rds"))

model_comp_logdiff_adult <- readRDS(here("cached_results/model_comp_logdiff_adult.Rds"))
model_comp_logdiff_preschooler <- readRDS(here("cached_results/model_comp_logdiff_preschooler.Rds"))
model_comp_logdiff_infant <- readRDS(here("cached_results/model_comp_logdiff_infant.Rds"))
```


# Model comparisons 

## Residual-based models 

```{r fi}
model_comp_logdiff_adult %>% select(-group) %>% 
  kable("latex", booktabs = TRUE, align = "c", 
        caption = "Adults") %>%
  kable_styling(full_width = TRUE, position = "left") %>%
  kable_styling(latex_options = "HOLD_position")
```


```{r}
model_comp_logdiff_preschooler %>% select(-group) %>% 
  kable("latex", booktabs = TRUE, align = "c", 
        caption = "Preschoolers") %>%
  kable_styling(full_width = TRUE, position = "left") %>%
  kable_styling(latex_options = "HOLD_position")
```


```{r}
model_comp_logdiff_infant %>% select(-group) %>% 
  kable("latex", booktabs = TRUE, align = "c", 
        caption = "Infants") %>%
  kable_styling(full_width = TRUE, position = "left") %>%
  kable_styling(latex_options = "HOLD_position")
```


## Robustness check models

```{r}
model_comp_logdiff_adult %>% select(-group) %>% 
  kable("latex", booktabs = TRUE, align = "c", 
        caption = "Adults") %>%
  kable_styling(full_width = TRUE, position = "left") %>%
  kable_styling(latex_options = "HOLD_position")
```


```{r}
model_comp_logdiff_preschooler %>% select(-group) %>% 
  kable("latex", booktabs = TRUE, align = "c", 
        caption = "Preschoolers") %>%
  kable_styling(full_width = TRUE, position = "left") %>%
  kable_styling(latex_options = "HOLD_position")

```


```{r}
model_comp_logdiff_infant %>% select(-group) %>% 
  kable("latex", booktabs = TRUE, align = "c", 
        caption = "Infants") %>%
  kable_styling(full_width = TRUE, position = "left") %>%
  kable_styling(latex_options = "HOLD_position")
```

# Correlations between measures and predictors 

Below are the correlation matrices for all predictors and the two operationalizations of dishabituation. Specifically, `resid_diff` represents the residual-based dishabituation, and `log_diff` is the difference score between the log-transformed looking time at the dishabituation trial and the last habituation trial. The numbers in the matrices are Pearson’s correlation coefficients, and blank cells indicate nonsignificant correlations.



```{r fig.cap="Correlational matrix for adults."}
corrplot(cormatrix_adult, 
         p.mat = pmatrix_adult, 
         method = "number",sig.level = 0.05, insig = "blank", title = "Adults", 
         mar = c(0, 0, 1, 0)) 
```


```{r fig.cap="Correlational matrix for preschoolers. "}
corrplot(cormatrix_preschooler, 
         p.mat = pmatrix_preschooler, 
         method = "number",sig.level = 0.05, insig = "blank", title = "Preschoolers", 
         mar = c(0, 0, 1, 0)) 
```



```{r fig.cap="Correlational matrix for infants."}
corrplot(cormatrix_infant, 
         p.mat = pmatrix_infant, 
         method = "number",sig.level = 0.05, insig = "blank", title = "Infants", 
         mar = c(0, 0, 1, 0)) 
```


