---
title: "Habituation re-analysis"
output:
  html_document:
    number_sections: false
    toc: true
    toc_float: true
---



```{r}
library(tidyverse)
library(here)
library(nlme)
library(lme4)
library(rugarch)
library(ggthemes)
library(car)

source("helper/volatility.R")
source("helper/slope.R")
```

Goal: analyzing the relationship between habituation patterns and the dishabituation magnitude across three datasets. 

- Adult dataset: https://onlinelibrary.wiley.com/doi/10.1111/tops.12631
- Preschooler dataset: https://escholarship.org/uc/item/5f38t3cd 
- Infant dataset: https://www.nature.com/articles/s41562-024-01965-x


# 1. Dataset prepping 

```{r}
adult_d <- read_csv(here("data/raw/adult_data.csv"))
preschooler_d <- read_csv(here("data/raw/preschooler_data.csv"))
infant_d <- read_csv(here("data/raw/infant_data.csv"))
```

## 1.1 Prep adult dataset 

Focusing on only the first block and blocks with more than two trials

```{r}
adult_habituation_data <- adult_d %>% 
  # first block
  filter(block_number == 1) %>% 
  # before the dishabituation trials
  filter(trial_number < deviant_position) %>% 
  # filtering out blocks with only two trials
  filter(deviant_position != 2) %>% 
  # create complexity note
  mutate(complexity = if_else(block_type == "complex_dissimilar", "complex", "simple")) %>% 
  # only keeping the relevant information 
  select(subject, trial_number, complexity, trial_looking_time) 


## add habituation residuals 

habituation_model <- nlme(
  trial_looking_time ~ a * exp(b * trial_number),
  data = adult_habituation_data,
  fixed = a + b ~ 1,                    # Fixed effects for decay
  random = b ~ 1 | subject,         # Random intercept and slope
  start = c(a = mean(adult_habituation_data$trial_looking_time), b = -0.1),        # Initial guesses for parameters
  control = nlmeControl(msMaxIter = 200, pnlsTol = 1e-6)  # Controls for stability
)

adult_habituation_data$resid_hab <- resid(habituation_model)

random_effects <- ranef(habituation_model)$b
fixed_decay_rate <- fixef(habituation_model)["b"]
adult_individual_decay_rates <- fixed_decay_rate + random_effects
adult_rate <- tibble("subject" = rownames(ranef(habituation_model)), 
                     "decay_rates" = adult_individual_decay_rates)


adult_habituation_data <- adult_habituation_data %>% 
  left_join(adult_rate, by = c("subject"))

```




## 1.2 Prep preschooler datset 

```{r}
# figure out what blocks to retain
retained_blocks <- preschooler_d %>% 
  filter(participant_type == "kids") %>% 
  distinct(subject, block_number, trial_number, trial_type) %>% 
  filter(trial_type == "deviant") %>% 
  filter(trial_number > 2) %>% 
  filter(block_number == 1) %>% 
  mutate(deviant_trial_number = trial_number)


preschooler_habituation_data <- preschooler_d %>% 
  # these are participants that have the first block meeting the criteria 
  filter(subject %in% retained_blocks$subject) %>% 
  filter(block_number == 1) %>% 
  left_join(retained_blocks %>% select(-trial_number, -trial_type), by = c("subject", "block_number")) %>% 
  filter(trial_number < deviant_trial_number) %>% 
  select(subject, trial_number, trial_looking_time, age_in_months) 

preschooler_habituation_data %>% 
  ungroup() %>% 
  distinct(subject, age_in_months) %>% 
  rename(age = age_in_months) %>% 
  write_csv(here("data/raw/preschooler_age.csv"))             

  habituation_model <- nlme(
  trial_looking_time ~ a * exp(b * trial_number),
  data = preschooler_habituation_data,
  fixed = a + b ~ 1,                    # Fixed effects for decay
  random = b ~ 1 | subject,         # Random intercept and slope
  start = c(a = mean(preschooler_habituation_data$trial_looking_time), b = -0.1),        # Initial guesses for parameters
  control = nlmeControl(msMaxIter = 200, pnlsTol = 1e-6)  # Controls for stability
)

preschooler_habituation_data$resid_hab <- resid(habituation_model)

random_effects <- ranef(habituation_model)$b
fixed_decay_rate <- fixef(habituation_model)["b"]
preschooler_individual_decay_rates <- fixed_decay_rate + random_effects

preschooler_rate <- tibble("subject" = rownames(ranef(habituation_model)), 
                     "decay_rates" = preschooler_individual_decay_rates)


preschooler_habituation_data <- preschooler_habituation_data %>% 
  left_join(preschooler_rate, by = c("subject"))
```


## 1.3 Prep infant dataset 

```{r}
infant_habituation_data <- infant_d %>% 
  select(paper, subj, specific_subject_id,agedays, starts_with("train")) %>% 
  pivot_longer(cols = starts_with("train")) %>% 
  filter(!is.na(value)) %>% 
  mutate(trial_number = str_replace(name, "train", "")) %>% 
  rename(trial_looking_time = value) %>% 
  rename(subject = specific_subject_id, age_in_days = agedays) %>% 
  select(paper, subject, trial_number, trial_looking_time, age_in_days) %>% 
  mutate(trial_number = as.numeric(trial_number))


infant_habituation_data %>% 
  distinct(subject, age_in_days) %>% 
  rename(age = age_in_days) %>% 
  write_csv(here("data/raw/infant_age.csv"))


infant_habituation_data


  

habituation_model <- nlme(
  trial_looking_time ~ a * exp(b * trial_number),
  data = infant_habituation_data,
  fixed = a + b ~ 1,                    # Fixed effects for decay
  random = b ~ 1 | subject,         # Random intercept and slope
  start = c(a = mean(infant_habituation_data$trial_looking_time), b = -0.1),        # Initial guesses for parameters
  control = nlmeControl(msMaxIter = 200, pnlsTol = 1e-6)  # Controls for stability
)

infant_habituation_data$resid_hab <- resid(habituation_model)

random_effects <- ranef(habituation_model)$b
fixed_decay_rate <- fixef(habituation_model)["b"]
infant_individual_decay_rates <- fixed_decay_rate + random_effects

infant_rate <- tibble("subject" = rownames(ranef(habituation_model)), 
                     "decay_rates" = infant_individual_decay_rates)


infant_habituation_data <- infant_habituation_data %>% 
  left_join(infant_rate, by = c("subject"))
```


# 2. Measure prepping 

## 2.1 Prep slope measure  

### 2.1.1 Demo how the curve change shape 

```{r}

# Define a function for the exponential decay
decay_model <- function(trial_number, a, b) {
  a * exp(b * trial_number)
}

# Create a data frame with different combinations of a and b
trial_data <- expand.grid(
  trial_number = seq(0, 10, by = 0.1),
  a = c(10, 20, 30),  # Different initial values
  b = c(0.05, -0.2, -0.6)  # Different decay rates
)

# Calculate the looking time based on the decay model
trial_data <- trial_data %>%
  mutate(trial_looking_time = decay_model(trial_number, a, b))

# Plot the curves
ggplot(trial_data, aes(x = trial_number, y = trial_looking_time, color = factor(a), linetype = factor(b))) +
  geom_line(size = 1) +
  labs(
    title = "Effect of Parameters 'a' and 'b' on Exponential Decay Curve",
    x = "Trial Number",
    y = "Trial Looking Time",
    color = "Initial Value (a)",
    linetype = "Decay Rate (b)"
  ) +
  theme_minimal()

```

### 2.1.2 Comparison on the model performance

```{r}

data <- adult_habituation_data

# get visualization to double check if things need to 
# simulation with confounding variables 


model_simple <- nlme(trial_looking_time ~ a * exp(b * trial_number),
                     data = data,
                     fixed = a + b ~ 1,
                     random = b ~ 1 | subject,
                     start = c(a = max(data$trial_looking_time), b = -0.05))

model_complex <- nlme(trial_looking_time ~ a * exp(b * trial_number),
                      data = data,
                      fixed = a + b ~ 1,
                      random =  a + b ~ 1 | subject,
                      start = c(a = max(data$trial_looking_time), b = -0.05))


model_simple_b <- nlme(trial_looking_time ~ a * exp(b * trial_number),
                     data = data,
                     fixed = a + b ~ 1,
                     random = a ~ 1 | subject,
                     start = c(a = max(data$trial_looking_time), b = -0.05))


random_effects <- ranef(model_complex)$b
fixed_decay_rate <- fixef(model_complex)["b"]
individual_decay_rates <- fixed_decay_rate + random_effects

random_effects <- ranef(model_simple)$b
fixed_decay_rate <- fixef(model_simple)["b"]
individual_decay_rates <- fixed_decay_rate + random_effects

anova(model_simple, model_complex) 
anova(model_simple_b, model_complex) # varying a
anova(model_simple_b, model_simple) 


anova(model_simple_b, model_simple) 


# a+b is comparable with a 
# a+b is worse than b
# a and b 

```


#### 2.1.2.1 model diagosnitic?


```{r}

model_simple <- nlme(trial_looking_time ~ a * exp(b * trial_number),
                     data = data,
                     fixed = a + b ~ 1,
                     random = b ~ 1 | subject,
                     start = c(a = max(data$trial_looking_time), b = -0.05))

data$simple_model_predict <- predict(model_simple)


# Select a random subset of subjects
set.seed(123) # For reproducibility
subset_subjects <- sample(unique(data$subject), 12) # Change 6 to the number of participants you want to plot

# Filter data for the selected subset
subset_data <- data %>% filter(subject %in% subset_subjects)


ggplot(subset_data, aes(x = trial_number)) +
  geom_point(aes(y = trial_looking_time), color = 'blue') +  # Actual data
  geom_line(aes(y = simple_model_predict), color = 'red', linetype = "dashed") +  # Model predictions
  labs(x = "Trial Number", y = "Looking Time", title = "Simple Model Fitting: Actual vs Predicted") +
  theme_minimal() + 
  facet_wrap(~subject)
```



```{r}
model_complex <- nlme(trial_looking_time ~ a * exp(b * trial_number),
                      data = data,
                      fixed = a + b ~ 1,
                      random =  a + b ~ 1 | subject,
                      start = c(a = max(data$trial_looking_time), b = -0.05))

data$complex_model_predict <- predict(model_complex)

# Select a random subset of subjects
set.seed(123) # For reproducibility
subset_subjects <- sample(unique(data$subject), 12) # Change 6 to the number of participants you want to plot
# Filter data for the selected subset
subset_data <- data %>% filter(subject %in% subset_subjects)

ggplot(subset_data, aes(x = trial_number)) +
  geom_point(aes(y = trial_looking_time), color = 'blue') +  # Actual data
  geom_line(aes(y = complex_model_predict), color = 'red', linetype = "dashed") +  # Model predictions
  labs(x = "Trial Number", y = "Looking Time", title = "Complex Model Fitting: Actual vs Predicted") +
  theme_minimal() + 
  facet_wrap(~subject)

```

```{r}
model_simple_b <- nlme(trial_looking_time ~ a * exp(b * trial_number),
                     data = data,
                     fixed = a + b ~ 1,
                     random = a ~ 1 | subject,
                     start = c(a = max(data$trial_looking_time), b = -0.05))

data$model_simple_b_predict <- predict(model_simple_b)

# Select a random subset of subjects
set.seed(123) # For reproducibility
subset_subjects <- sample(unique(data$subject), 12) # Change 6 to the number of participants you want to plot
# Filter data for the selected subset
subset_data <- data %>% filter(subject %in% subset_subjects)

ggplot(subset_data, aes(x = trial_number)) +
  geom_point(aes(y = trial_looking_time), color = 'blue') +  # Actual data
  geom_line(aes(y = model_simple_b_predict), color = 'red', linetype = "dashed") +  # Model predictions
  labs(x = "Trial Number", y = "Looking Time", title = "Simple Model B Fitting: Actual vs Predicted") +
  theme_minimal() + 
  facet_wrap(~subject)
```





## 2.2 Prep volatility measure 


```{r}
adult_volatility_measure <- get_volatility_measures(adult_habituation_data)
preschooler_volatility_measure <-  get_volatility_measures(preschooler_habituation_data)
infant_volatility_measure <-  get_volatility_measures(infant_habituation_data)
```



# 3. Dishabituation and measure dataset prepping 

## 3.1 Adult 

### 3.1.1 get difference 

```{r}

adult_dishab_raw <- adult_d %>% 
  filter(block_number == 1) %>% 
  filter(trial_number == deviant_position) %>% 
  filter(deviant_position != 2) 


deviant_model <- lme(
  trial_looking_time ~ 1,  # Deviant trial modeled as intercept-only
  random = ~ 1 | subject,  # Random intercept to account for subject variability
  data = adult_dishab_raw
)

adult_dishab_raw$resid_dev <- resid(deviant_model)


adult_dishab_d <- adult_dishab_raw %>% 
  mutate(complexity = if_else(block_type == "complex_dissimilar", "complex", "simple")) %>% 
  mutate(dishab_trial_num = trial_number + 1) %>% 
   select(subject, resid_dev, dishab_trial_num, complexity)

adult_complexity <- adult_dishab_raw %>% 
  mutate(complexity = if_else(block_type == "complex_dissimilar", "complex", "simple")) %>% 
  select(subject,  complexity)

write_csv(adult_complexity, here("data/raw/adult_complexity.csv"))

```



### 3.1.2 link age 


```{r}
raw_adult <- read_csv(here("data/raw/adult_raw_with_prolific.csv"))
prolific_d <- read_csv(here("data/raw/adult_prolific.csv"))

subject_prolific_link <- raw_adult %>% 
  filter(grepl("p_id", responses)) %>% 
  mutate(prolific_id = str_extract(responses, '(?<="p_id":")[^"]+')) %>% 
  select(subject, prolific_id)

link_df <- prolific_d %>% 
  rename(prolific_id = participant_id) %>% 
  select(prolific_id, age) %>% 
  left_join(subject_prolific_link, by = c("prolific_id"))

link_df %>% 
  select(age, subject) %>% 
  filter(!is.na(subject)) %>% 
  filter(subject %in% adult_complexity$subject) %>% 
  write_csv(here("data/raw/adult_age.csv"))



adult_dishab_d <- adult_dishab_d %>% 
  left_join(link_df %>% select(-prolific_id), by = c("subject")) 
```


### 3.1.3 combine with the measures 

```{r}
adult_measures_with_dishab <- adult_dishab_d %>% 
  left_join(adult_volatility_measure, by = c("subject")) %>% 
  left_join(adult_habituation_data %>% group_by(subject) %>% filter(trial_number == max(trial_number)) %>% rename(last_resid_hab = resid_hab) %>% distinct(subject, last_resid_hab, decay_rates)) %>% 
  mutate(resid_diff = resid_dev - last_resid_hab)
```


## 3.2 Preschooler 

### 3.2.1 Put together 

```{r}





preschooler_dishab_raw <- preschooler_d %>% 
  filter(child_age_group != "adult") %>% 
  filter(block_number == 1) %>% 
  filter(subject %in% retained_blocks$subject) %>% 
  left_join(retained_blocks %>% select(-trial_number, -trial_type), by = c("subject", "block_number")) %>% 
  filter(trial_number == deviant_trial_number) 


deviant_model <- lme(
  trial_looking_time ~ 1,  # Deviant trial modeled as intercept-only
  random = ~ 1 | subject,  # Random intercept to account for subject variability
  data = preschooler_dishab_raw
)

preschooler_dishab_raw$resid_dev <- resid(deviant_model)

preschooler_dishab_d <- preschooler_dishab_raw %>% 
  mutate(complexity = if_else(block_type == "complex_dissimilar", "complex", "simple")) %>% 
  mutate(dishab_trial_num = trial_number + 1) %>% 
   select(subject, resid_dev, dishab_trial_num, complexity)

preschooler_measures_with_dishab <- preschooler_dishab_d %>% 
  left_join(preschooler_volatility_measure, by = c("subject")) %>% 
  left_join(preschooler_habituation_data %>% group_by(subject) %>% filter(trial_number == max(trial_number)) %>% rename(last_resid_hab = resid_hab) %>% distinct(subject, last_resid_hab, decay_rates)) %>% 
  mutate(resid_diff = resid_dev - last_resid_hab)
```


## 3.3 Infants 

### 3.1.1 Re-wrangle some datasets 

```{r}
# try to construct the dataset to mimic the other ones, using only first of the unexpected as the "dishabituating trials"
# figure out the trial number for the extra trial
last_trial_number <- infant_d %>% 
  rename(subject = specific_subject_id) %>% 
  select(subject, agedays, starts_with("train"), "expected1") %>% 
  rename(trainEXTRA = expected1) %>% 
  pivot_longer(cols = starts_with("train")) %>% 
  filter(!is.na(value)) %>% 
  mutate(trial_number = as.numeric(str_replace(name, "train", ""))) %>% 
 group_by(subject) %>% 
  summarise(max_trial = max(trial_number, na.rm = TRUE)) %>% 
  mutate(last_trial = max_trial + 1) %>% 
  select(subject,last_trial) %>% 
  ungroup() %>% 
  filter(!last_trial == -Inf)

```


```{r}

infant_model_d <- infant_d %>% 
  mutate(complexity = if_else(domain == "psychology", "complex", "simple")) %>% 
  rename(subject = specific_subject_id) %>% 
  select(subject, agedays, complexity, starts_with("train"), "expected1") %>% 
  rename(trainEXTRA = expected1) %>% 
  pivot_longer(cols = starts_with("train")) %>% 
  filter(!is.na(value)) %>% 
  mutate(trial_number = as.numeric(str_replace(name, "train", ""))) %>% 
  left_join(last_trial_number, by = c("subject")) %>% 
  mutate(trial_type = case_when(
    is.na(trial_number) ~ "deviant", 
    TRUE ~ "background"
  )) %>% 
  mutate(trial_number = case_when(
    is.na(trial_number) ~ last_trial, 
    TRUE ~ trial_number
  )) %>% 
  filter(!is.na(trial_number)) %>% 
  select(subject, trial_number, trial_type, value, agedays, complexity) %>% 
  rename(trial_looking_time = value)

```




### 3.1.3  Put things together 

```{r}
duplicate_subjects <- infant_model_d %>% group_by(subject) %>% filter(trial_type != "deviant") %>%  filter(trial_number == max(trial_number)) %>% select(subject, trial_number, trial_looking_time, agedays) %>% group_by(subject) %>% count() %>% filter(n > 1)

duplicate_subjects %>% write_csv(here("data/raw/infant_bad.csv"))
```


```{r}

infant_d %>% 
  rename(subject = specific_subject_id) %>% 
  select(subject, domain) %>% 
  mutate(complexity = if_else(domain == "psychology", "complex", "simple")) %>% 
  distinct(subject, complexity) %>% 
  filter(!subject %in% duplicate_subjects) %>% 
  write_csv(here("data/raw/infant_complexity.csv"))



infant_dishab_raw <- infant_model_d %>% 
  filter(trial_type == "deviant") %>% 
  filter(!subject %in% duplicate_subjects$subject) 


deviant_model <- lme(
  trial_looking_time ~ 1,  # Deviant trial modeled as intercept-only
  random = ~ 1 | subject,  # Random intercept to account for subject variability
  data = infant_dishab_raw
)

infant_dishab_raw$resid_dev <- resid(deviant_model)

infant_dishab_raw_dishab_d <- infant_dishab_raw %>% 
  mutate(dishab_trial_num = trial_number) %>% 
  select(subject, resid_dev, dishab_trial_num, complexity, agedays)

infant_measures_with_dishab <- infant_dishab_raw_dishab_d %>% 
  left_join(infant_volatility_measure, by = c("subject")) %>% 
  left_join(infant_habituation_data %>% group_by(subject) %>% filter(trial_number == max(trial_number)) %>% rename(last_resid_hab = resid_hab) %>% distinct(subject, last_resid_hab, decay_rates)) %>% 
  mutate(resid_diff = resid_dev - last_resid_hab)

```



# 4. Visualization 

## 4.1. Raw habituation data Visualizatino 

### 4.1.1 Adults

```{r}
adult_habituation_data %>% 
  ggplot(aes(x = trial_number, y = trial_looking_time, group = subject)) + 
  geom_point(alpha = .1) + 
  geom_line(alpha = .1) 
 

adult_habituation_data %>% 
  ggplot(aes(x = trial_number, y = trial_looking_time, group = subject)) + 
  geom_point(alpha = .1) + 
  geom_line(alpha = .1) +
  scale_y_log10()

adult_habituation_data %>% 
  ggplot(aes(x = trial_number, y = trial_looking_time)) + 
  geom_point(alpha = .1) + 
  stat_summary(fun.data = "mean_cl_boot", geom = "pointrange", color = "red") 
 

adult_habituation_data %>% 
  ggplot(aes(x = trial_number, y = trial_looking_time)) + 
  geom_point(alpha = .1) + 
  stat_summary(fun.data = "mean_cl_boot", geom = "pointrange", color = "red") +
  scale_y_log10()
```



### 4.1.2 Preschoolers

```{r}
preschooler_habituation_data %>% 
  ggplot(aes(x = trial_number, y = trial_looking_time, group = subject)) + 
  geom_point(alpha = .1) + 
  geom_line(alpha = .1) 
 

preschooler_habituation_data %>% 
  ggplot(aes(x = trial_number, y = trial_looking_time, group = subject)) + 
  geom_point(alpha = .1) + 
  geom_line(alpha = .1) +
  scale_y_log10()

preschooler_habituation_data %>% 
  ggplot(aes(x = trial_number, y = trial_looking_time)) + 
  geom_point(alpha = .1) + 
  stat_summary(fun.data = "mean_cl_boot", geom = "pointrange", color = "red") 
 

preschooler_habituation_data %>% 
  ggplot(aes(x = trial_number, y = trial_looking_time)) + 
  geom_point(alpha = .1) + 
  stat_summary(fun.data = "mean_cl_boot", geom = "pointrange", color = "red") +
  scale_y_log10()
```



### 4.1.3 Infants

```{r}
infant_habituation_data %>% 
  ggplot(aes(x = trial_number, y = trial_looking_time, group = subject)) + 
  geom_point(alpha = .1) + 
  geom_line(alpha = .1) 
 

infant_habituation_data %>% 
  ggplot(aes(x = trial_number, y = trial_looking_time, group = subject))  + 
  geom_point(alpha = .1) + 
  geom_line(alpha = .1) +
  scale_y_log10()

infant_habituation_data %>% 
  ggplot(aes(x = trial_number, y = trial_looking_time)) + 
  geom_point(alpha = .1) + 
  stat_summary(fun.data = "mean_cl_boot", geom = "pointrange", color = "red") 
 

infant_habituation_data %>% 
  ggplot(aes(x = trial_number, y = trial_looking_time)) + 
  geom_point(alpha = .1) + 
  stat_summary(fun.data = "mean_cl_boot", geom = "pointrange", color = "red") +
  scale_y_log10()
```



# 4.2 Raw dishabituation data visualization 

```{r}
adult_measures_with_dishab %>% 
  ggplot(aes(x = dishab_diff)) + 
  geom_density(alpha = .1, aes(fill = as.factor(dishab_trial_num)))
```

```{r}
preschooler_measures_with_dishab %>% 
  ggplot(aes(x = dishab_diff)) + 
  geom_density(alpha = .1, aes(fill = as.factor(dishab_trial_num)))
```

```{r}
infant_measure_with_dishab %>% 
  ggplot(aes(x = dishab_diff)) + 
  geom_density()
```




# 4.3 Measure and data predictions 


```{r}
full_d <- bind_rows(
  infant_measures_with_dishab %>% mutate(group = "infants"), 
  preschooler_measures_with_dishab %>% mutate(group = "preschoolers"), 
  adult_measures_with_dishab %>% mutate(group = "adults")
)
```

### 4.3.1 Slope measures vs resid across age groups 

```{r}
full_d %>% 
  ggplot(aes(x = decay_rates, y = resid_diff)) + 
  geom_point(alpha = .2) + 
  facet_wrap(~group, scales = "free") + 
  geom_smooth(method = "lm") 


full_d %>% 
  ggplot(aes(x = decay_rates, y = resid_diff)) + 
  geom_point(alpha = .2) + 
  facet_wrap(complexity~group, scales = "free") + 
  geom_smooth(method = "lm") 


```

## 4.3.2 Volatility measurs across age groups 



```{r fig.width=8, fig.height=8}
full_d %>% 
  pivot_longer(cols = ends_with("volatility"), names_to = "volatility_type", values_to =  "volatility_measure") %>% 
  ggplot(aes(x = volatility_measure, y = resid_dev)) + 
  geom_point(alpha = .2) + 
  facet_wrap(volatility_type~group, scales = "free") + 
  geom_smooth(method = "lm") 
```


# 5. Models 

## 5.1 Adults 

### 5.1.2 Decay Rate only 

```{r}
lm(dishab_diff ~ decay_rate, data = adult_measures_with_dishab) %>% summary()


lm(dishab_diff ~ complexity, data = adult_measures_with_dishab) %>% summary()

```


### 5.1.3 Decay Rate + Age 

```{r}
lm(dishab_diff ~ decay_rate + age, data = adult_measures_with_dishab) %>% summary()
```




### 5.1.3 Decay Rate + Age + volatility measure

```{r}
lm(dishab_diff ~ decay_rate + age + sd_volatility, data = adult_measures_with_dishab) %>% summary()
```

## 5.2 Preschoolers 

### 5.1.2 Decay Rate only 

```{r}
lm(dishab_diff ~ decay_rate , data = preschooler_measures_with_dishab) %>% summary()
```



### 5.1.3 Decay Rate + Age 

```{r}
lm(dishab_diff ~ decay_rate + age_in_months, data = preschooler_measures_with_dishab) %>% summary()
```



### 5.1.3 Decay Rate + Age + volatility measure

```{r}
lm(dishab_diff ~ decay_rate + age_in_months + sd_volatility, data = preschooler_measures_with_dishab) %>% summary()
```


```{r}
vif(lm(dishab_diff ~ decay_rate + age_in_months + sd_volatility, data = preschooler_measures_with_dishab))
```


## 6.1 Infants 

### 6.1.1 Decay Rate only 

```{r}
lm(dishab_diff ~ decay_rate , data = infant_measure_with_dishab) %>% summary()
```


### 5.1.3 Decay Rate + Age 

```{r}
lm(dishab_diff ~ decay_rate + agedays, data = infant_measure_with_dishab) %>% summary()
```

### 5.1.3 Decay Rate + Age + volatility measure

```{r}
lm(dishab_diff ~ decay_rate + agedays + sd_volatility, data = infant_measure_with_dishab) %>% summary()
```



# 6. Summarizing Model results

## 6.1 Combine Decay Rate Only models

```{r}

bind_rows(
  lm(dishab_diff ~ decay_rate, data = adult_measures_with_dishab) %>% broom::tidy(conf.int = TRUE, conf.level = 0.95) %>% mutate(group = "adults"),
  lm(dishab_diff ~ decay_rate, data = preschooler_measures_with_dishab) %>% broom::tidy(conf.int = TRUE, conf.level = 0.95) %>% mutate(group = "preschoolers"),
  lm(dishab_diff ~ decay_rate, data = infant_measure_with_dishab) %>% broom::tidy(conf.int = TRUE, conf.level = 0.95) %>% mutate(group = "infants")
) %>% 
  filter(term == "decay_rate") %>% 
  ggplot(aes(x = factor(group, levels = c("infants", "preschoolers", "adults")), y = estimate, ymin = conf.low, ymax = conf.high)) + 
  geom_pointrange() + 
  geom_hline(yintercept = 0, color = "gray", linetype = "dashed") + 
  coord_flip() + 
  theme_few() + 
  xlab("")


```


## 6.2 Combine Decay Rate + Age Models

```{r}
bind_rows(
   lm(dishab_diff ~ decay_rate + age, data = adult_measures_with_dishab) %>% broom::tidy(conf.int = TRUE, conf.level = 0.95) %>% mutate(group = "adults"),
  lm(dishab_diff ~ decay_rate + age_in_months, data = preschooler_measures_with_dishab) %>% broom::tidy(conf.int = TRUE, conf.level = 0.95) %>% mutate(group = "preschoolers"),
  lm(dishab_diff ~ decay_rate + agedays, data = infant_measure_with_dishab) %>% broom::tidy(conf.int = TRUE, conf.level = 0.95) %>% mutate(group = "infants")
) %>% 
  filter(term != "(Intercept)") %>% 
  mutate(clean_term = case_when(
    term == "age_in_months" ~ "age",
    term == "agedays" ~ "age", 
    TRUE ~ term
)) %>% 
  ggplot(aes(x = factor(group, levels = c("infants", "preschoolers", "adults")), y = estimate, ymin = conf.low, ymax = conf.high)) + 
  geom_pointrange() + 
  geom_hline(yintercept = 0, color = "gray", linetype = "dashed") + 
  coord_flip() + 
  theme_few() + 
  xlab("") + 
  facet_wrap(~clean_term)
```



## 6.3 Combine Decay Rate + Age + Volatility Models

```{r}
bind_rows(
   lm(dishab_diff ~ decay_rate + age + sd_volatility, data = adult_measures_with_dishab) %>% broom::tidy(conf.int = TRUE, conf.level = 0.95) %>% mutate(group = "adults"),
  lm(dishab_diff ~ decay_rate + age_in_months + sd_volatility, data = preschooler_measures_with_dishab) %>% broom::tidy(conf.int = TRUE, conf.level = 0.95) %>% mutate(group = "preschoolers"),
  lm(dishab_diff ~ decay_rate + agedays + sd_volatility, data = infant_measure_with_dishab) %>% broom::tidy(conf.int = TRUE, conf.level = 0.95) %>% mutate(group = "infants")
) %>% 
  filter(term != "(Intercept)") %>% 
  mutate(clean_term = case_when(
    term == "age_in_months" ~ "age",
    term == "agedays" ~ "age", 
    TRUE ~ term
)) %>% 
  ggplot(aes(x = factor(group, levels = c("infants", "preschoolers", "adults")), y = estimate, ymin = conf.low, ymax = conf.high)) + 
  geom_pointrange() + 
  geom_hline(yintercept = 0, color = "gray", linetype = "dashed") + 
  coord_flip() + 
  theme_few() + 
  xlab("") + 
  facet_wrap(~clean_term, scales = "free")
```

```{r}
bind_rows(
   lm(dishab_diff ~ decay_rates_random + age + sd_volatility, data = adult_measures_with_dishab) %>% broom::tidy(conf.int = TRUE, conf.level = 0.95) %>% mutate(group = "adults"),
  lm(dishab_diff ~ decay_rates_random + age_in_months + sd_volatility, data = preschooler_measures_with_dishab) %>% broom::tidy(conf.int = TRUE, conf.level = 0.95) %>% mutate(group = "preschoolers"),
  lm(dishab_diff ~ decay_rates_random + agedays + sd_volatility, data = infant_measure_with_dishab) %>% broom::tidy(conf.int = TRUE, conf.level = 0.95) %>% mutate(group = "infants")
) %>% 
  filter(term != "(Intercept)") %>% 
  mutate(clean_term = case_when(
    term == "age_in_months" ~ "age",
    term == "agedays" ~ "age", 
    TRUE ~ term
)) %>% 
  ggplot(aes(x = factor(group, levels = c("infants", "preschoolers", "adults")), y = estimate, ymin = conf.low, ymax = conf.high)) + 
  geom_pointrange() + 
  geom_hline(yintercept = 0, color = "gray", linetype = "dashed") + 
  coord_flip() + 
  theme_few() + 
  xlab("") + 
  facet_wrap(~clean_term, scales = "free")
```


