



```{r}
library(tidyverse)
library(here)
library(lme4)
library(nlme)

# load simulation data 
sim_dr <- read_csv(here("data/simulations/sim_dr.csv"))
sim_ft <- read_csv(here("data/simulations/sim_ft.csv"))
sim_nr <- read_csv(here("data/simulations/sim_nr.csv"))

# load real data
adult_hab_full <- read_csv(here("data/raw/adult_hab_full.csv"))
adult_hab_no_last <- read_csv(here("data/raw/adult_hab_no_last.csv"))
adult_dishab <- read_csv(here("data/raw/adult_dishab.csv"))

preschooler_hab_full <- read_csv(here("data/raw/preschooler_hab_full.csv"))
preschooler_hab_no_last <- read_csv(here("data/raw/preschooler_hab_no_last.csv"))
preschooler_dishab <- read_csv(here("data/raw/preschooler_dishab.csv"))

infant_hab_full <- read_csv(here("data/raw/infant_hab_full.csv"))
infant_hab_no_last <- read_csv(here("data/raw/infant_hab_no_last.csv"))
infant_dishab <- read_csv(here("data/raw/infant_dishab.csv"))

# helper function 
source(here("helper/slope.R"))
source(here("helper/resid.R"))
```



# 0. Preprocessing 

## 0.1. DR

```{r}
sim_dr_hab <-  sim_dr %>% 
  filter(trial_number < deviant_position) %>% 
  select(subject, trial_number, trial_looking_time) 

sim_dr_hab_no_last <-  sim_dr %>% 
  filter(trial_number < deviant_position - 1) %>% 
  select(subject, trial_number, trial_looking_time) 

sim_dr_dev <-  sim_dr %>% 
  filter(trial_number == deviant_position) %>% 
  select(subject, trial_number, trial_looking_time) 

```

## 0.2 FT 

```{r}
sim_ft_hab <-  sim_ft %>% 
  filter(trial_number < deviant_position) %>% 
  select(subject, trial_number, trial_looking_time) 

sim_ft_hab_no_last <-  sim_ft %>% 
  filter(trial_number < deviant_position - 1) %>% 
  select(subject, trial_number, trial_looking_time) 

sim_ft_dev <-  sim_ft %>% 
  filter(trial_number == deviant_position) %>% 
  select(subject, trial_number, trial_looking_time) 
```

## 0.3 NR 

```{r}
sim_nr_hab <-  sim_nr %>% 
  filter(trial_number < deviant_position) %>% 
  select(subject, trial_number, trial_looking_time) 

sim_nr_hab_no_last <-  sim_nr %>% 
  filter(trial_number < deviant_position - 1) %>% 
  select(subject, trial_number, trial_looking_time) 

sim_nr_dev <-  sim_nr %>% 
  filter(trial_number == deviant_position) %>% 
  select(subject, trial_number, trial_looking_time) 
```


# 1. Slope INCLUDE last hab trial - Resid Deviant Models 

## 1.1 Simulations 

### 1.1.1 NR

```{r}

## get slope 
sim_nr_slope <- get_individual_slope(sim_nr_hab)

## get residual 

sim_nr_hab_resid <- get_hab_residual(sim_nr_hab)
#sim_nr_dev_resid <- get_dev_residual(sim_nr_dev)

sim_nr_slope %>% 
  left_join(sim_nr_dev_resid, by = c("subject")) %>% 
  ggplot(aes(x = decay_rate, y = resid_dev)) + 
  geom_point() + 
  geom_smooth(method = "lm")

```


### 1.1.2 FT

```{r}
## get slope 
sim_ft_slope <- get_individual_slope(sim_ft_hab)

## get residual 

sim_ft_hab_resid <- get_hab_residual(sim_ft_hab)
#sim_ft_dev_resid <- get_dev_residual(sim_ft_dev)

sim_ft_slope %>% 
  left_join(sim_ft_dev_resid, by = c("subject")) %>% 
  ggplot(aes(x = decay_rate, y = resid_dev)) + 
  geom_point() + 
  geom_smooth(method = "lm")
```

### 1.1.3 DR

```{r}
## get slope 
#sim_dr_slope <- get_individual_slope(sim_dr_hab)

## get residual 

#sim_dr_hab_resid <- get_hab_residual(sim_dr_hab)
#sim_dr_dev_resid <- get_dev_residual(sim_dr_dev)

sim_dr_slope %>% 
  left_join(sim_dr_dev_resid, by = c("subject")) %>% 
  ggplot(aes(x = decay_rate, y = resid_dev)) + 
  geom_point() + 
  geom_smooth(method = "lm")
```

# 1.2 Real Data


```{r}

hab_datasets <- list(adult_hab_full, preschooler_hab_full, infant_hab_full)  
dev_datasets <- list(adult_dishab, preschooler_dishab, infant_dishab)  

# Get slopes for all hab datasets
slopes <- map(hab_datasets, get_individual_slope)

# Get hab residuals for all hab datasets
hab_residuals <- map(hab_datasets, get_hab_residual)

# Get dev residuals for all dev datasets
dev_residuals <- map(dev_datasets, get_dev_residual)

# Left join for each pair of datasets 
joined_results <- map2(slopes, dev_residuals, ~ left_join(.x, .y, by = "subject"))

bind_rows(
  joined_results[[1]] %>% mutate(group = "1_adult"), 
  joined_results[[2]] %>% mutate(group = "2_preschooler"),
  joined_results[[3]] %>% mutate(group = "3_infant"),
) %>% 
  ggplot(aes(x = decay_rate, y = resid_dev)) + 
  geom_point() + 
  geom_smooth(method = "lm") + 
  facet_wrap(~group, scales = "free")

```

# 2. Slope INCLUDE last hab trial - (Resid Deviant - Resid Last Trial) 

## 2.1 Simulations 

### 2.1.1 NR

```{r}

# get slope 
sim_nr_slope <- get_individual_slope(sim_nr_hab)

# get residual 
sim_nr_hab_resid <- get_hab_residual(sim_nr_hab)
sim_nr_dev_resid <- get_dev_residual(sim_nr_dev)

# calculate the diff between the last trial and the deviant
hab_dev_diff <- sim_nr_hab_resid %>% 
  group_by(subject) %>% 
  filter(trial_number == max(trial_number)) %>% 
  rename(last_hab_resid = resid_hab) %>% 
  select(subject, last_hab_resid) %>% 
  left_join(
    sim_nr_dev_resid %>% select(subject, resid_dev), 
    by = c("subject")
  ) %>% 
  mutate(resid_diff = resid_dev - last_hab_resid)



hab_dev_diff %>% 
  left_join(sim_nr_slope, by = c("subject")) %>% 
  ggplot(aes(x = decay_rate, y = resid_diff)) + 
  geom_point() + 
  geom_smooth(method = "lm")

```


### 2.1.2 FT

```{r}
## get slope 
sim_ft_slope <- get_individual_slope(sim_ft_hab)

## get residual 
sim_ft_hab_resid <- get_hab_residual(sim_ft_hab)
sim_ft_dev_resid <- get_dev_residual(sim_ft_dev)

# calculate the diff between the last trial and the deviant
hab_dev_diff <- sim_ft_hab_resid %>% 
  group_by(subject) %>% 
  filter(trial_number == max(trial_number)) %>% 
  rename(last_hab_resid = resid_hab) %>% 
  select(subject, last_hab_resid) %>% 
  left_join(
    sim_ft_dev_resid %>% select(subject, resid_dev), 
    by = c("subject")
  ) %>% 
  mutate(resid_diff = resid_dev - last_hab_resid)



hab_dev_diff %>% 
  left_join(sim_ft_slope, by = c("subject")) %>% 
  ggplot(aes(x = decay_rate, y = resid_diff)) + 
  geom_point() + 
  geom_smooth(method = "lm")
```

### 2.1.3 DR

```{r}
## get slope 
sim_dr_slope <- get_individual_slope(sim_dr_hab)

## get residual 
sim_dr_hab_resid <- get_hab_residual(sim_dr_hab)
sim_dr_dev_resid <- get_dev_residual(sim_dr_dev)

# calculate the diff between the last trial and the deviant
hab_dev_diff <- sim_dr_hab_resid %>% 
  group_by(subject) %>% 
  filter(trial_number == max(trial_number)) %>% 
  rename(last_hab_resid = resid_hab) %>% 
  select(subject, last_hab_resid) %>% 
  left_join(
    sim_dr_dev_resid %>% select(subject, resid_dev), 
    by = c("subject")
  ) %>% 
  mutate(resid_diff = resid_dev - last_hab_resid)



hab_dev_diff %>% 
  left_join(sim_dr_slope, by = c("subject")) %>% 
  ggplot(aes(x = decay_rate, y = resid_diff)) + 
  geom_point() + 
  geom_smooth(method = "lm")
```

## 2.2 Real Data 

```{r}
calculate_resid_diff <- function(hab_resid, dev_resid){
  
  resid_diff <- hab_resid %>% ha
    group_by(subject) %>% 
    filter(trial_number == max(trial_number)) %>% 
    rename(last_hab_resid = resid_hab) %>% 
   select(subject, last_hab_resid) %>% 
    left_join(
    dev_resid %>% select(subject, resid_dev), 
    by = c("subject")
  ) %>% 
  mutate(resid_diff = resid_dev - last_hab_resid)
  
  return(resid_diff)
}

# calculate the residual difference 
resid_diffs <- map2(hab_residuals, dev_residuals, calculate_resid_diff)

# join the datasets 
joined_results <- map2(slopes, resid_diffs, ~ left_join(.x, .y, by = "subject"))

# visualize 
bind_rows(
  joined_results[[1]] %>% mutate(group = "1_adult"), 
  joined_results[[2]] %>% mutate(group = "2_preschooler"),
  joined_results[[3]] %>% mutate(group = "3_infant"),
) %>% 
  ggplot(aes(x = decay_rate, y = resid_diff)) + 
  geom_point() + 
  geom_smooth(method = "lm") + 
  facet_wrap(~group, scales = "free")


```


# 3. Slope BEFORE last habituation trial - (Raw Dev - Raw Last)

## 3.1 Simulations 

### 3.1.1 NR

```{r}
# get slope 
sim_nr_slope <- get_individual_slope(sim_nr_hab_no_last)


# calculate the diff between the last trial and the deviant
hab_dev_diff <- sim_nr %>% 
  filter(trial_type != "deviant") %>% 
  group_by(subject) %>% 
  filter(trial_number == max(trial_number)) %>% 
  rename(last_hab_looking_time = trial_looking_time) %>% 
  select(subject, last_hab_looking_time) %>% 
  left_join(
    sim_nr %>% filter(trial_type == "deviant") %>% 
    rename(dev_looking_time = trial_looking_time) %>% 
    select(subject, dev_looking_time), 
    by = c("subject")
  ) %>% 
  mutate(
    hab_dev_diff = dev_looking_time - last_hab_looking_time, 
    log_hab_dev_diff = log(dev_looking_time)- log(last_hab_looking_time))



hab_dev_diff %>% 
  left_join(sim_nr_slope, by = c("subject")) %>% 
  ggplot(aes(x = decay_rate, y = hab_dev_diff)) + 
  geom_point() + 
  geom_smooth(method = "lm")

hab_dev_diff %>% 
  left_join(sim_nr_slope, by = c("subject")) %>% 
  ggplot(aes(x = decay_rate, y = log_hab_dev_diff)) + 
  geom_point() + 
  geom_smooth(method = "lm")

```

### 3.1.2 FT

```{r}
# get slope 
sim_ft_slope <- get_individual_slope(sim_ft_hab_no_last)


# calculate the diff between the last trial and the deviant
hab_dev_diff <- sim_ft %>% 
  filter(trial_type != "deviant") %>% 
  group_by(subject) %>% 
  filter(trial_number == max(trial_number)) %>% 
  rename(last_hab_looking_time = trial_looking_time) %>% 
  select(subject, last_hab_looking_time) %>% 
  left_join(
    sim_ft %>% filter(trial_type == "deviant") %>% 
    rename(dev_looking_time = trial_looking_time) %>% 
    select(subject, dev_looking_time), 
    by = c("subject")
  ) %>% 
  mutate(
    hab_dev_diff = dev_looking_time - last_hab_looking_time, 
    log_hab_dev_diff = log(dev_looking_time)- log(last_hab_looking_time))



hab_dev_diff %>% 
  left_join(sim_ft_slope, by = c("subject")) %>% 
  ggplot(aes(x = decay_rate, y = hab_dev_diff)) + 
  geom_point() + 
  geom_smooth(method = "lm")

hab_dev_diff %>% 
  left_join(sim_ft_slope, by = c("subject")) %>% 
  ggplot(aes(x = decay_rate, y = log_hab_dev_diff)) + 
  geom_point() + 
  geom_smooth(method = "lm")
```

### 3.1.3 DR

```{r}
# get slope 
sim_dr_slope <- get_individual_slope(sim_dr_hab_no_last)


# calculate the diff between the last trial and the deviant
hab_dev_diff <- sim_dr %>% 
  filter(trial_type != "deviant") %>% 
  group_by(subject) %>% 
  filter(trial_number == max(trial_number)) %>% 
  rename(last_hab_looking_time = trial_looking_time) %>% 
  select(subject, last_hab_looking_time) %>% 
  left_join(
    sim_dr %>% filter(trial_type == "deviant") %>% 
    rename(dev_looking_time = trial_looking_time) %>% 
    select(subject, dev_looking_time), 
    by = c("subject")
  ) %>% 
  mutate(
    hab_dev_diff = dev_looking_time - last_hab_looking_time, 
    log_hab_dev_diff = log(dev_looking_time)- log(last_hab_looking_time))



hab_dev_diff %>% 
  left_join(sim_dr_slope, by = c("subject")) %>% 
  ggplot(aes(x = decay_rate, y = hab_dev_diff)) + 
  geom_point() + 
  geom_smooth(method = "lm")

hab_dev_diff %>% 
  left_join(sim_dr_slope, by = c("subject")) %>% 
  ggplot(aes(x = decay_rate, y = log_hab_dev_diff)) + 
  geom_point() + 
  geom_smooth(method = "lm")
```


## 3.2 Real Data 

```{r}

get_diff_raw_looking_time <- function(hab_data, dev_data){
  # calculate the diff between the last trial and the deviant
  hab_dev_diff <- hab_data %>% 
  group_by(subject) %>% 
  filter(trial_number == max(trial_number)) %>% 
  rename(last_hab_looking_time = trial_looking_time) %>% 
  select(subject, last_hab_looking_time) %>% 
  left_join(
    dev_data  %>% 
    rename(dev_looking_time = trial_looking_time) %>% 
    select(subject, dev_looking_time), 
    by = c("subject")
  ) %>% 
  mutate(
    hab_dev_diff = dev_looking_time - last_hab_looking_time, 
    log_hab_dev_diff = log(dev_looking_time)- log(last_hab_looking_time))
  
  return(hab_dev_diff)
}


hab_no_last_datasets <- list(adult_hab_no_last, preschooler_hab_no_last, infant_hab_no_last)  
hab_full_datasets <- list(adult_hab_full, preschooler_hab_full, infant_hab_full)
dev_datasets <- list(adult_dishab, preschooler_dishab, infant_dishab)

# Get slopes for all hab datasets
slopes <- map(hab_no_last_datasets, get_individual_slope)
diff_raw <- map2(hab_full_datasets, dev_datasets, get_diff_raw_looking_time)


# Left join for each pair of datasets 
joined_results <- map2(slopes, diff_raw, ~ left_join(.x, .y, by = "subject"))

bind_rows(
  joined_results[[1]] %>% mutate(group = "1_adult"), 
  joined_results[[2]] %>% mutate(group = "2_preschooler"),
  joined_results[[3]] %>% mutate(group = "3_infant"),
) %>% 
  ggplot(aes(x = decay_rate, y = hab_dev_diff)) + 
  geom_point() + 
  geom_smooth(method = "lm") + 
  facet_wrap(~group, scales = "free")


bind_rows(
  joined_results[[1]] %>% mutate(group = "1_adult"), 
  joined_results[[2]] %>% mutate(group = "2_preschooler"),
  joined_results[[3]] %>% mutate(group = "3_infant"),
) %>% 
  ggplot(aes(x = decay_rate, y = log_hab_dev_diff)) + 
  geom_point() + 
  geom_smooth(method = "lm") + 
  facet_wrap(~group, scales = "free")
```

# 4. slope INCLUDE last hab trial - RAW deviant - Residual last trial 


## 4.1 Simulations 

### 4.1.1 NR 

```{r}
# get slope 
sim_nr_slope <- get_individual_slope(sim_nr_hab)

# get residual 
sim_nr_hab_resid <- get_hab_residual(sim_nr_hab)

hab_dev_diff <- sim_nr_hab_resid %>% 
  group_by(subject) %>% 
  filter(trial_number == max(trial_number)) %>% 
  rename(last_hab_resid = resid_hab) %>% 
  select(subject, last_hab_resid) %>% 
  left_join(
    sim_nr_dev %>% 
    rename(dev_looking_time = trial_looking_time) %>% 
    select(subject, dev_looking_time), 
    by = c("subject")
  ) %>% 
  mutate(
    hab_resid_dev_diff = dev_looking_time - last_hab_resid)

hab_dev_diff %>% 
  left_join(sim_nr_slope, by = c("subject")) %>% 
  ggplot(aes(x = decay_rate, y = hab_resid_dev_diff)) + 
  geom_point() + 
  geom_smooth(method = "lm")


```

### 4.1.2 FT

```{r}
# get slope 
sim_ft_slope <- get_individual_slope(sim_ft_hab)

# get residual 
sim_ft_hab_resid <- get_hab_residual(sim_ft_hab)

hab_dev_diff <- sim_ft_hab_resid %>% 
  group_by(subject) %>% 
  filter(trial_number == max(trial_number)) %>% 
  rename(last_hab_resid = resid_hab) %>% 
  select(subject, last_hab_resid) %>% 
  left_join(
    sim_ft_dev %>% 
    rename(dev_looking_time = trial_looking_time) %>% 
    select(subject, dev_looking_time), 
    by = c("subject")
  ) %>% 
  mutate(
    hab_resid_dev_diff = dev_looking_time - last_hab_resid)

hab_dev_diff %>% 
  left_join(sim_ft_slope, by = c("subject")) %>% 
  ggplot(aes(x = decay_rate, y = hab_resid_dev_diff)) + 
  geom_point() + 
  geom_smooth(method = "lm")

```

### 4.1.3 DR 

```{r}
# get slope 
sim_dr_slope <- get_individual_slope(sim_dr_hab)

# get residual 
sim_dr_hab_resid <- get_hab_residual(sim_dr_hab)

hab_dev_diff <- sim_ft_hab_resid %>% 
  group_by(subject) %>% 
  filter(trial_number == max(trial_number)) %>% 
  rename(last_hab_resid = resid_hab) %>% 
  select(subject, last_hab_resid) %>% 
  left_join(
    sim_dr_dev %>% 
    rename(dev_looking_time = trial_looking_time) %>% 
    select(subject, dev_looking_time), 
    by = c("subject")
  ) %>% 
  mutate(
    hab_resid_dev_diff = dev_looking_time - last_hab_resid)

hab_dev_diff %>% 
  left_join(sim_dr_slope, by = c("subject")) %>% 
  ggplot(aes(x = decay_rate, y = hab_resid_dev_diff)) + 
  geom_point() + 
  geom_smooth(method = "lm")
```

## 4.2 Real Data

```{r}
hab_full_datasets <- list(adult_hab_full, preschooler_hab_full, infant_hab_full)
dev_datasets <- list(adult_dishab, preschooler_dishab, infant_dishab)

get_diff_resid_looking_time <- function(hab_resid_data, dev_data){
  # calculate the diff between the last trial and the deviant
  hab_dev_diff <- hab_resid_data %>% 
  group_by(subject) %>% 
  filter(trial_number == max(trial_number)) %>% 
  rename(last_hab_resid = resid_hab) %>% 
  select(subject, last_hab_resid) %>% 
  left_join(
    dev_data  %>% 
    rename(dev_looking_time = trial_looking_time) %>% 
    select(subject, dev_looking_time), 
    by = c("subject")
  ) %>% 
  mutate(
    hab_dev_diff = dev_looking_time - last_hab_resid)
  
  return(hab_dev_diff)
}


# Get slopes for all hab datasets
slopes <- map(hab_full_datasets, get_individual_slope)

# Get residuals for all datasets 
hab_residuals <- map(hab_full_datasets, get_hab_residual)

#calculate the difference 
diff_resid_lt <- map2(hab_residuals, dev_datasets, get_diff_resid_looking_time)

#join the dataset 
joined_results <- map2(slopes, diff_resid_lt, ~ left_join(.x, .y, by = "subject"))

#visualization 
# visualize 
bind_rows(
  joined_results[[1]] %>% mutate(group = "1_adult"), 
  joined_results[[2]] %>% mutate(group = "2_preschooler"),
  joined_results[[3]] %>% mutate(group = "3_infant"),
) %>% 
  ggplot(aes(x = decay_rate, y = hab_dev_diff)) + 
  geom_point() + 
  geom_smooth(method = "lm") + 
  facet_wrap(~group, scales = "free")

```


# 5. Checking 

```{r}

#https://www.statforbiology.com/2019/stat_nlmm_designconstraints/

sim_ft_hab$subject <- as.factor(sim_ft_hab$subject)

mixed_exp_model <- nlme(
    trial_looking_time ~ a * exp(b * trial_number),  # Simplified exponential decay model
    data = sim_ft_hab,
    fixed = a + b ~ 1,                # Fixed effects for a and b
    random = list(subject = pdDiag(a + b ~ 1)),       # Random effect for b (decay rate) by subject
    
    start = c(a = max(sim_ft_hab$trial_looking_time), 
              b = -0.05)  # Starting values
  )



 mixed_exp_model <- nlme(
    trial_looking_time ~ a * exp(b * trial_number),  # Simplified exponential decay model
    data = infant_hab_no_last,
    fixed = a + b ~ 1,                # Fixed effects for a and b
     random = list(subject = pdDiag(a + b ~ 1)),         # Random effect for b (decay rate) by subject
    start = c(a = max(infant_hab_no_last$trial_looking_time), 
              b = -0.05)  # Starting values
  )


fixef(mixed_exp_model)

coefficients(mixed_exp_model)

VarCorr(mixed_exp_model)



as.data.frame(coefficients(mixed_exp_model)) %>% 
  ggplot(aes(x = a, y = b)) + 
  geom_point() + 
  geom_smooth(method = "lm") +
  labs(title = "a and b")

  
cor(coefficients(mixed_exp_model)$a, coefficients(mixed_exp_model)$b)

sim_ft_hab %>% 
  filter(trial_number == 1) %>% 
  mutate(
    b_val = ranef(mixed_exp_model)$b
  ) %>% 
  ggplot(aes(x = trial_looking_time, y = b_val)) + 
  geom_point() + 
  geom_smooth(method = "lm") + 
  labs(title = "first trial and b")

sim_ft_hab %>% 
  filter(trial_number == 1) %>% 
  mutate(
    a_val = ranef(mixed_exp_model)$a
  ) %>% 
  ggplot(aes(x = trial_looking_time, y = a_val)) + 
  geom_point() + 
  geom_smooth(method = "lm")+
  labs(title = "first trial and a")


```


```{r}
data_summary <- sim_ft_hab %>%
  group_by(subject) %>%
  summarize(
    a = max(trial_looking_time),
    b = lm(log(trial_looking_time) ~ trial_number)$coefficients[2]
  )


data_summary %>% 
  ggplot(aes(x = a, y = b)) + 
  geom_point() + 
  geom_smooth(method = "lm")

```


```{r}

library(dplyr)

n_subjects  <- 50
n_trials    <- 20
a_range     <- c(5000, 7000)  # narrower range for 'a'
b_range     <- c(-0.08, -0.03) # narrower range for 'b' (decay rate)
residual_sd <- 150             # moderate residual noise

sim_data <- lapply(seq_len(n_subjects), function(i) {
  # Draw subject-specific true parameters:
  a_i <- runif(1, a_range[1], a_range[2])
  b_i <- runif(1, b_range[1], b_range[2])

  # Generate times
  trial_number <- seq_len(n_trials)

  # True expected looking time
  true_looking <- a_i * exp(b_i * trial_number)

  # Add some random noise
  observed_looking <- true_looking + rnorm(n_trials, mean = 0, sd = residual_sd)

  # Enforce a minimum > 0 (sometimes large negative residual can produce small or negative values)
  observed_looking <- pmax(observed_looking, 10)

  # Return a data frame for this subject
  data.frame(
    subject      = factor(i),
    trial_number = trial_number,
    looking_time = observed_looking,
    true_a       = a_i,
    true_b       = b_i
  )
}) %>%
  bind_rows()

start_vals <- c(a = mean(a_range), b = mean(b_range))

fit_nlme <- nlme(
  trial_looking_time ~ a * exp(b * trial_number),
  data   = adult_hab_no_last,
  fixed  = a + b ~ 1,
  random = list(subject = pdDiag(a + b ~ 1)),  # random effect only on b
  start  = start_vals,
  control = nlmeControl(
    pnlsMaxIter = 50,
    msMaxIter   = 200,
    tolerance   = 1e-6,
    opt         = "nlminb"
  )
)

fixed_effects <- fixef(fit_nlme)

fitted_random_effects <- ranef(fit_nlme) %>% 
  rownames_to_column("subject") %>% 
  mutate(a = a + fixed_effects['a'], 
         b = b + fixed_effects['b'])



comparison <- left_join(fitted_random_effects, 
                        adult_hab_no_last %>%
                          group_by(subject) %>%
                          summarize(true_a = first(true_a), true_b = first(true_b)),
                                                                                               by = "subject")

# Plot for 'a'
ggplot(comparison, aes(x = true_a, y = a)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0, color = "blue") +
  geom_smooth(method = "lm") + 
  labs(title = "True vs Fitted Random Effects for a",
       x = "True Random Effect (a)", y = "Fitted Random Effect (a)")

ggplot(comparison, aes(x = true_a, y = true_b)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0, color = "blue") +
  labs(title = "True A vs True B",
       x = "True Random Effect (a)", y = "True Random Effect (b)")

# Plot for 'b'
ggplot(comparison, aes(x = true_b, y = b)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0, color = "blue") +
  geom_smooth(method = "lm") + 
  labs(title = "True vs Fitted Random Effects for b",
       x = "True Random Effect (b)", y = "Fitted Random Effect (b)")


ggplot(fitted_random_effects, aes(x = a, y = b)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, color = "blue") +
  labs(title = "Random Effects for a vs b",
       x = "Random Effect for a", y = "Random Effect for b")

summary(simulated_data)

```


```{r}
library(brms)
set.seed(123)
n_subjects <- 30
n_trials <- 15

# Ranges for 'a' and 'b'
a_range <- c(5000, 10000)
b_range <- c(-0.1, -0.01)
residual_sd <- 50

simulated_data <- do.call(rbind, lapply(seq_len(n_subjects), function(i) {
  a_true <- runif(1, a_range[1], a_range[2])
  b_true <- runif(1, b_range[1], b_range[2])
  trial_number <- 1:n_trials
  trial_looking_time <- a_true * exp(b_true * trial_number) + 
    rnorm(n_trials, 0, residual_sd)
  data.frame(
    subject = factor(paste0("S", sprintf("%03d", i))),
    trial_number = trial_number,
    trial_looking_time = trial_looking_time,
    true_a = a_true,
    true_b = b_true
  )
}))

nl_formula <- bf(
  trial_looking_time ~ a * exp(b * trial_number),  # The equation
  a + b ~ 1 + (1|subject),                         # Random intercepts for 'a' and 'b'
  nl = TRUE
)

priors <- c(
  prior(normal(7500, 2000), nlpar = "a", lb = 0),     # a is presumably large and positive
  prior(normal(-0.05, 0.02), nlpar = "b")            # b is negative
)

bayesian_model <- brm(
  formula = nl_formula,
  data = simulated_data,
  family = gaussian(),   # We'll assume Gaussian residuals
  prior = priors,
  chains = 4,            # Number of MCMC chains
  cores = 4,             # Number of CPU cores for parallel
  iter = 8000,           # Total iterations per chain
  warmup = 1000,         # Burn-in/warm-up
  control = list(adapt_delta = 0.99),  # More conservative adaptation
  seed = 123
)
```

```{r}
ranef(bayesian_model)$subject %>% as.data.frame() %>% 
  ggplot(aes(x = Estimate.a_Intercept, y = Estimate.b_Intercept)) + 
  geom_point() + 
  geom_smooth(method = "lm")
```

