
```{r}
library(tidyverse)
library(here)
library(ggthemes)

full_d <- read_csv(here("data/processed.csv")) %>% 
  # dealing with some age formatting 
  mutate(
    age = case_when(
      !is.na(agedays) ~ agedays, 
      !is.na(age_in_months) ~ age_in_months,
      TRUE ~ age
    )
  )


```


```{r}

# TODO: corplot of all measures 
cor((full_d %>% 
  filter(group == "adults"))$sd_volatility, 
  (full_d %>% 
  filter(group == "adults"))$mad_robust_volatility, 
  use = "complete.obs")
```

# 1. Construct all formula 

```{r}

get_formulas <- function(){
  volatility_meausures <- c("sd_volatility", 
                            "mad_volatility", 
                            "cv_volatility",
                            "range_volatility",
                            "mad_robust_volatility",
                            "garch_volatility"
                            )
  formulas <- lapply(volatility_meausures, function(x){
    paste0(
      "dishab_diff ~ age + complexity + decay_rate + ", 
      x
      
    )
  })
  
  formula_res <- tibble(
    volatility_meausures, 
    formulas
  ) %>% 
    unnest(c(formulas)) %>% 
    mutate(
      formula_no_complexity = gsub("\\s*\\+\\s*complexity", "", formulas))

  
  return(formula_res)
  
}

formulas <- get_formulas()
formulas
```
# 2. Run all the models

```{r}

nested_d <- full_d %>% 
  group_by(group) %>% 
  nest()



safe_lm <- function(formula, data) {
  tryCatch({
    # Attempt to fit the lmer model
    lm(as.formula(formula), data = data)
  }, error = function(e) {
    # Return NULL or error message when an error occurs
    message(paste("Error in fitting model:", e$message))
    NULL
  })
}


all_models <- nested_d %>%
  crossing(formulas) %>%  
  mutate(
    model = case_when(
      group != "preschoolers" ~ map2(data, formulas, ~ safe_lm(.y, .x)),
      group == "preschoolers" ~ map2(data, formula_no_complexity, ~ safe_lm(.y, .x))
    
  )) %>% 
  mutate(
    tidy_results = map(model, ~ if (!is.null(.)) broom::tidy(., conf.int = TRUE, conf.level = 0.95) else NULL)) 


```


# 3. Model comparions 


```{r}
all_models %>% 
  select(group, volatility_meausures, model) %>% 
  mutate(
    AIC = map(model, ~ if (!is.null(.)) AIC(.) else NULL)
  ) %>% 
  unnest(AIC) %>% 
  group_by(group) %>% 
  arrange(AIC)
```

# 4. Check model results 

```{r}
all_models %>% 
  filter(volatility_meausures == "cv_volatility") %>% 
  unnest(tidy_results) %>% 
  filter(term != "(Intercept)") %>% 
  ggplot(aes(x = factor(group, levels = c("infants", "preschoolers", "adults")), y = estimate, ymin = conf.low, ymax = conf.high)) + 
  geom_pointrange() + 
  geom_hline(yintercept = 0, color = "gray", linetype = "dashed") + 
  coord_flip() + 
  theme_few() + 
  xlab("") + 
  facet_wrap(~term, scales = "free")
```

```{r}
(all_models %>% 
  filter(volatility_meausures == "sd_volatility") %>% 
  filter(group == "preschoolers"))$model[[1]] %>% 
  summary()
```
```{r}
(all_models %>% 
  filter(volatility_meausures == "sd_volatility") %>% 
  filter(group == "infants"))$model[[1]] %>% 
  summary()
```

```{r}
(all_models %>% 
  filter(volatility_meausures == "sd_volatility") %>% 
  filter(group == "adults"))$model[[1]] %>% 
  summary()
```

Figures: 
- participants, tasks -> description of the dataset
- demo of the slope and intercept, volatility to give ppl conceptual understanding of what we meant 
- results plot - raw data? model data?? 

todo: 
- VIF model checking? 



Check with Mike: 
- model varying a and b? 
- volatiltiy measure 
- get more data to estimate 






