---
title: "Habituation re-analysis"
output:
  html_document:
    code_folding: hide
    number_sections: false
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  warning = TRUE,   # Hide warnings
  message = FALSE    # Hide messages
)
```

```{r}
library(tidyverse)
library(here)
library(lme4)
library(nlme)

# load real data
adult_hab_no_last <- read_csv(here("data/raw/adult_hab_no_last.csv"))
preschooler_hab_no_last <- read_csv(here("data/raw/preschooler_hab_full.csv"))
infant_hab_no_last <- read_csv(here("data/raw/infant_hab_no_last.csv"))
```

# TLDR:

| **Age Group**    | **Single-Subject (NLS)**                                   | **NLME**<br>*`pdDiag(a + b ~ 1)`* | **NLME**<br>*`pdSymm(a + b ~ 1)`* |
|:----------------:|:---------------------------------------------------------:|:--------------------------------:|:---------------------------------:|
| **Adults**       | Negative correlation<br>(\(\hat{a}_i\) vs. \(\hat{b}_i\)) | Positive correlation             | Fails to converge                 |
| **Preschoolers** | Negative correlation<br>(\(\hat{a}_i\) vs. \(\hat{b}_i\)) | Positive correlation             | Fails to converge                 |
| **Infants**      | Negative correlation<br>(\(\hat{a}_i\) vs. \(\hat{b}_i\)) | No correlation                   | Negative correlation              |


# 0. Set up Model Demo

```{r}
fit_nlme <- function(data, start_vals){
  fit_nlme <- nlme(
  trial_looking_time ~ a * exp(b * trial_number),
  data   = data,
  fixed  = a + b ~ 1,
  random = list(subject = pdDiag(a + b ~ 1)),  # random effect only on b
  start  = start_vals,
  control = nlmeControl(
    pnlsMaxIter = 50,
    msMaxIter   = 200,
    tolerance   = 1e-6,
    opt         = "nlminb"
  )
)
} 

fit_nlme_pdSymm <- function(data, start_vals){
  fit_nlme <- nlme(
  trial_looking_time ~ a * exp(b * trial_number),
  data   = data,
  fixed  = a + b ~ 1,
  random = list(subject = pdSymm(a + b ~ 1)),  # random effect only on b
  start  = start_vals,
  control = nlmeControl(
    pnlsMaxIter = 50,
    msMaxIter   = 200,
    tolerance   = 1e-6,
    opt         = "nlminb"
  )
)
} 

fit_nlme_b_only <- function(data, start_vals){
  fit_nlme <- nlme(
  trial_looking_time ~ a * exp(b * trial_number),
  data   = data,
  fixed  = a + b ~ 1,
  random = b ~ 1 | subject,  # random effect only on b
  start  = start_vals,
  control = nlmeControl(
    pnlsMaxIter = 50,
    msMaxIter   = 200,
    tolerance   = 1e-6,
    opt         = "nlminb"
  )
)
} 

fit_nlme_a_only <- function(data, start_vals){
  fit_nlme <- nlme(
  trial_looking_time ~ a * exp(b * trial_number),
  data   = data,
  fixed  = a + b ~ 1,
  random = a ~ 1 | subject,  # random effect only on b
  start  = start_vals,
  control = nlmeControl(
    pnlsMaxIter = 50,
    msMaxIter   = 200,
    tolerance   = 1e-6,
    opt         = "nlminb"
  )
)
} 
```


# 1. Generate Fake Dataset where we know a and b

## 1.1 Generate the data 


```{r}
n_subjects  <- 50
n_trials    <- 20
a_range     <- c(5000, 7000)  # narrower range for 'a'
b_range     <- c(-0.08, -0.03) # narrower range for 'b' (decay rate)
residual_sd <- 150             # moderate residual noise

sim_data <- lapply(seq_len(n_subjects), function(i) {
  # Draw subject-specific true parameters:
  a_i <- runif(1, a_range[1], a_range[2])
  b_i <- runif(1, b_range[1], b_range[2])

  # Generate times
  trial_number <- seq_len(n_trials)

  # True expected looking time
  true_looking <- a_i * exp(b_i * trial_number)

  # Add some random noise
  observed_looking <- true_looking + rnorm(n_trials, mean = 0, sd = residual_sd)

  # Enforce a minimum > 0 (sometimes large negative residual can produce small or negative values)
  observed_looking <- pmax(observed_looking, 10)

  # Return a data frame for this subject
  data.frame(
    subject      = factor(i),
    trial_number = trial_number,
    trial_looking_time = observed_looking,
    true_a       = a_i,
    true_b       = b_i
  )
}) %>%
  bind_rows()
```

## 1.2 Show real a and b independence 

```{r}
sim_data %>% 
  distinct(true_a, true_b) %>% 
  ggplot(aes(x = true_a, y = true_b)) + 
  geom_point() + 
  geom_smooth(method = "lm")
```


## 1.3 Show model demonstration 

```{r}
start_vals <- c(a = mean(a_range), b = mean(b_range))
sim_model <- fit_nlme(sim_data, start_vals)

fixed_effects <- fixef(sim_model)

fitted_random_effects <- ranef(sim_model) %>% 
  rownames_to_column("subject") %>% 
  mutate(a = a + fixed_effects['a'], 
         b = b + fixed_effects['b'])



comparison <- left_join(fitted_random_effects, 
                        sim_data %>%
                          group_by(subject) %>%
                          summarize(true_a = first(true_a), true_b = first(true_b)),
                                                                                               by = "subject")

# Plot for 'a'
ggplot(comparison, aes(x = true_a, y = a)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0, color = "blue") +
  geom_smooth(method = "lm") + 
  labs(title = "True vs Fitted Random Effects for a",
       x = "True Random Effect (a)", y = "Fitted Random Effect (a)")

ggplot(comparison, aes(x = true_a, y = true_b)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0, color = "blue") +
  geom_smooth(method = "lm") + 
  labs(title = "True A vs True B",
       x = "True Random Effect (a)", y = "True Random Effect (b)")

# Plot for 'b'
ggplot(comparison, aes(x = true_b, y = b)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0, color = "blue") +
  geom_smooth(method = "lm") + 
  labs(title = "True vs Fitted Random Effects for b",
       x = "True Random Effect (b)", y = "Fitted Random Effect (b)")


ggplot(fitted_random_effects, aes(x = a, y = b)) +
  geom_point() +
  geom_smooth(method = "lm") + 
  labs(title = "Random Effects for a vs b",
       x = "Random Effect for a", y = "Random Effect for b")
```


# 2. On Real Data with pdDiag

## 2.1 Adults 

```{r}
adult_start_vals <- c(a = mean(adult_hab_no_last$trial_looking_time), b = -0.05)

adult_model <- fit_nlme(adult_hab_no_last, adult_start_vals)
adult_model_b_only <- fit_nlme_b_only(adult_hab_no_last, adult_start_vals)
adult_model_a_only <- fit_nlme_a_only(adult_hab_no_last, adult_start_vals)


anova.gls(adult_model, adult_model_b_only)
anova(adult_model, adult_model_a_only)
anova(adult_model_a_only, adult_model_b_only)

VarCorr(adult_model)
VarCorr(adult_model_a_only)
VarCorr(adult_model_b_only)


AIC_full <- AIC(adult_model)
AIC_b_only <- AIC(adult_model_b_only)
AIC_a_only <- AIC(adult_model_a_only)

BIC_full <- BIC(adult_model)
BIC_b_only <- BIC(adult_model_b_only)
BIC_a_only <- BIC(adult_model_a_only)

# Combine into a table
model_comparison <- data.frame(
  Model = c("Random a + b", "Random b only", "Random a only"),
  AIC = c(AIC_full, AIC_b_only, AIC_a_only),
  BIC = c(BIC_full, BIC_b_only, BIC_a_only)
)

model_comparison

fixed_effects <- fixef(adult_model)

fitted_random_effects <- ranef(adult_model) %>% 
  rownames_to_column("subject") %>% 
  mutate(a = a + fixed_effects['a'], 
         b = b + fixed_effects['b'])

ggplot(fitted_random_effects, aes(x = a, y = b)) +
  geom_point() +
  geom_smooth(method = "lm") + 
  labs(title = "Random Effects for a vs b",
       x = "Random Effect for a", y = "Random Effect for b")
```



## 2.2 Preschooler

```{r}
preschooler_start_vals <- c(a = mean(preschooler_hab_no_last$trial_looking_time), b = -0.05)

preschooler_model <- fit_nlme(preschooler_hab_no_last, preschooler_start_vals)
fixed_effects <- fixef(preschooler_model)
preschooler_model_b_only <- fit_nlme_b_only(preschooler_hab_no_last, preschooler_start_vals)
preschooler_model_a_only <- fit_nlme_a_only(preschooler_hab_no_last, preschooler_start_vals)



anova(preschooler_model, preschooler_model_b_only)
anova(preschooler_model, preschooler_model_a_only)
anova(preschooler_model_a_only, preschooler_model_b_only)

VarCorr(preschooler_model)
VarCorr(preschooler_model_a_only)
VarCorr(preschooler_model_b_only)


AIC_full <- AIC(preschooler_model)
AIC_b_only <- AIC(preschooler_model_b_only)
AIC_a_only <- AIC(preschooler_model_a_only)

BIC_full <- BIC(preschooler_model)
BIC_b_only <- BIC(preschooler_model_b_only)
BIC_a_only <- BIC(preschooler_model_a_only)

# Combine into a table
model_comparison <- data.frame(
  Model = c("Random a + b", "Random b only", "Random a only"),
  AIC = c(AIC_full, AIC_b_only, AIC_a_only),
  BIC = c(BIC_full, BIC_b_only, BIC_a_only)
)

fitted_random_effects <- ranef(preschooler_model) %>% 
  rownames_to_column("subject") %>% 
  mutate(a = a + fixed_effects['a'], 
         b = b + fixed_effects['b'])

ggplot(fitted_random_effects, aes(x = a, y = b)) +
  geom_point() +
  geom_smooth(method = "lm") + 
  labs(title = "Random Effects for a vs b",
       x = "Random Effect for a", y = "Random Effect for b")
```

## 2.3 On Infant data 

```{r}
infant_start_vals <- c(a = mean(infant_hab_no_last$trial_looking_time), b = -0.05)

infant_model <- fit_nlme(infant_hab_no_last, infant_start_vals)
fixed_effects <- fixef(infant_model)
infant_model_b_only <- fit_nlme_b_only(infant_hab_no_last, infant_start_vals)
infant_model_a_only <- fit_nlme_a_only(infant_hab_no_last, infant_start_vals)



anova(infant_model, infant_model_b_only)
anova(infant_model, infant_model_a_only)
anova(infant_model_a_only, infant_model_b_only)

VarCorr(infant_model)
VarCorr(infant_model_a_only)
VarCorr(infant_model_b_only)


AIC_full <- AIC(infant_model)
AIC_b_only <- AIC(infant_model_b_only)
AIC_a_only <- AIC(infant_model_a_only)

BIC_full <- BIC(infant_model)
BIC_b_only <- BIC(infant_model_b_only)
BIC_a_only <- BIC(infant_model_a_only)

# Combine into a table
model_comparison <- data.frame(
  Model = c("Random a + b", "Random b only", "Random a only"),
  AIC = c(AIC_full, AIC_b_only, AIC_a_only),
  BIC = c(BIC_full, BIC_b_only, BIC_a_only)
)

fitted_random_effects <- ranef(infant_model) %>% 
  rownames_to_column("subject") %>% 
  mutate(a = a + fixed_effects['a'], 
         b = b + fixed_effects['b'])

ggplot(fitted_random_effects, aes(x = a, y = b)) +
  geom_point() +
  geom_smooth(method = "lm") + 
  labs(title = "Random Effects for a vs b",
       x = "Random Effect for a", y = "Random Effect for b")
```


# 3. On Real Data with pdSymm

## 3.1 Adults 

Model can't converge 

```{r}
adult_start_vals <- c(a = mean(adult_hab_no_last$trial_looking_time), b = -0.05)

# Can't converge 
#adult_model <- fit_nlme_pdSymm(adult_hab_no_last, adult_start_vals)


```


## 3.2 Preschooler

False Convergence 

```{r}
preschooler_start_vals <- c(a = mean(preschooler_hab_no_last$trial_looking_time), b = -0.05)


preschooler_model <- fit_nlme_pdSymm(preschooler_hab_no_last, preschooler_start_vals)
fixed_effects <- fixef(preschooler_model)

fitted_random_effects <- ranef(preschooler_model) %>% 
  rownames_to_column("subject") %>% 
  mutate(a = a + fixed_effects['a'], 
         b = b + fixed_effects['b'])

ggplot(fitted_random_effects, aes(x = a, y = b)) +
  geom_point() +
  geom_smooth(method = "lm") + 
  labs(title = "Random Effects for a vs b",
       x = "Random Effect for a", y = "Random Effect for b")
```

## 3.3 On Infant data

false convergence (8)

```{r}
infant_start_vals <- c(a = mean(infant_hab_no_last$trial_looking_time), b = -0.05)

infant_model <- fit_nlme_pdSymm(infant_hab_no_last, infant_start_vals)
fixed_effects <- fixef(infant_model)

fitted_random_effects <- ranef(infant_model) %>% 
  rownames_to_column("subject") %>% 
  mutate(a = a + fixed_effects['a'], 
         b = b + fixed_effects['b'])

ggplot(fitted_random_effects, aes(x = a, y = b)) +
  geom_point() +
  geom_smooth(method = "lm") + 
  labs(title = "Random Effects for a vs b",
       x = "Random Effect for a", y = "Random Effect for b")
```



# 4. On Real Data but NLS 

```{r}
fit_single_subject_exp <- function(df) {

  start_vals <- list(
    a = max(df$trial_looking_time, na.rm = TRUE),
    b = -0.05
  )


  nls_fit <- tryCatch(
    nls(
      trial_looking_time ~ a * exp(b * trial_number),
      data    = df,
      start   = start_vals,
      control = nls.control(maxiter = 200)
    ),
    error = function(e) NULL
  )
  
  # If the fit failed, return NA; otherwise return the estimated coefficients
  if (is.null(nls_fit)) {
    return(data.frame(a = NA_real_, b = NA_real_))
  } else {
    coefs <- coef(nls_fit)
    return(data.frame(a = coefs["a"], b = coefs["b"]))
  }
}
```

## 4.1 Adults 

```{r}
adult_subject_estimates <- adult_hab_no_last %>%
  group_by(subject) %>%
  do(fit_single_subject_exp(.)) %>%
  ungroup()

cor(adult_subject_estimates$a, adult_subject_estimates$b, use = "complete.obs")

adult_subject_estimates %>% 
  ggplot(aes(x = a, y = b)) + 
  geom_point() + 
  geom_smooth(method = "lm")
```

## 4.2 Preschooler 

```{r}
preschooler_subject_estimates <- preschooler_hab_no_last %>%
  group_by(subject) %>%
  do(fit_single_subject_exp(.)) %>%
  ungroup()

cor(preschooler_subject_estimates$a, preschooler_subject_estimates$b, use = "complete.obs")

preschooler_subject_estimates %>% 
  ggplot(aes(x = a, y = b)) + 
  geom_point() + 
  geom_smooth(method = "lm")
```

## 4.3 Infants 

```{r}


infant_subject_estimates <- infant_hab_no_last %>%
  group_by(subject) %>%
  do(fit_single_subject_exp(.)) %>%
  ungroup()

cor(infant_subject_estimates$a, infant_subject_estimates$b, use = "complete.obs")

infant_subject_estimates %>% 
  ggplot(aes(x = a, y = b)) + 
  geom_point() + 
  geom_smooth(method = "lm")


```

