```{r}
library(tidyverse)
library(here)
library(ggthemes)

d <- read_csv("processed_rt_task_data.csv")
```

# Slope 

First let's just focus on all first blocks? 

```{r}
slope_data <- d %>% 
  filter(block_number == 1) %>% 
  filter(trial_number < deviant_position) %>% 
  filter(deviant_position != 2) %>% 
  select(subject, trial_number, trial_looking_time) 
```


OK, first trying to fit individual exponential slope, but too much non-converging? 

```{r}
# Load necessary packages
library(dplyr)

# Initialize a list to store the slopes for each subject
slopes <- data.frame(subject = unique(slope_data$subject), decay_rate = NA)

# Loop over each subject
for (subj in unique(slope_data$subject)) {
  # Subset data for each subject
  subject_data <- slope_data %>% filter(subject == subj)
  
  # Fit the exponential model using nls
  tryCatch({
    # Adjust starting values based on data characteristics
    fit <- nls(trial_looking_time ~ a * exp(b * trial_number) + c, 
               data = subject_data,
               start = list(a = max(subject_data$trial_looking_time), 
                            b = -0.05, 
                            c = min(subject_data$trial_looking_time)))
    
    # Extract the decay rate (slope)
    decay_rate <- coef(fit)["b"]
    
    # Store the result in the slopes data frame
    slopes$decay_rate[slopes$subject == subj] <- decay_rate
  }, error = function(e) {
    message(paste("Model failed to converge for subject:", subj))
  })
}

# View the slopes for each subject
print(slopes)

slopes %>% filter(!is.na(decay_rate))
```

now let's try the mixed model approach 


```{r}

library(lme4)

# Adjust starting values if the model has trouble converging
start_a <- max(slope_data$trial_looking_time)
start_b <- -0.05 
start_c <- min(slope_data$trial_looking_time)

# Fit the nonlinear mixed-effects model
mixed_exp_model <- nlmer(
  trial_looking_time ~ a * exp(b * trial_number) + c ~ (b | subject),  # Formula for exponential decay with random slopes on b
  data = slope_data,
  start = c(a = max(slope_data$trial_looking_time), 
            b = -0.05, 
            c = min(slope_data$trial_looking_time)) 
)

mixed_exp_model <- nlme(
  trial_looking_time ~ a * exp(b * trial_number),  # Simplified exponential decay model
  data = slope_data,
  fixed = a + b ~ 1,                # Fixed effects for a and b
  random = b ~ 1 | subject,         # Random effect for b (decay rate) by subject
  start = c(a = max(slope_data$trial_looking_time), b = -0.05)  # Starting values
)

# Summary of the model to view the fixed and random effects
summary(mixed_exp_model)

# Extract random effects (individual decay rates)
random_effects <- ranef(mixed_exp_model)$b

# Combine fixed effect with random effect for individual slopes
fixed_decay_rate <- fixef(mixed_exp_model)["b"]
individual_decay_rates <- fixed_decay_rate + random_effects

# Create a data frame to store the slopes for each participant
slopes <- data.frame(
  subject = rownames(ranef(mixed_exp_model)),
  decay_rate = individual_decay_rates
)

# View the individual slopes for each participant
print(slopes)


```


OK yay this works!! now let's see if it predicts anything about the dishabituation?!?!?!

```{r}

test_d <- d %>% 
  filter(block_number == 1) %>% 
  filter(trial_number == deviant_position) %>% 
  filter(deviant_position != 2) %>% 
  select(subject, block_type, trial_looking_time) %>% 
  left_join(slopes, by = c("subject")) 


test_d %>% 
  ggplot(aes(x = decay_rate, y = trial_looking_time)) + 
  geom_point(aes(color = block_type)) + 
  theme_few() 
  
lm(trial_looking_time ~ decay_rate + block_type , data = test_d) %>% 
  summary()


```


but doesn't control for general sloweness of the person????


what if we do something like controlling using the average time 

- high collinearity, also no more significant value :| 


```{r}

mean_lt_d <- d %>% 
  filter(block_number == 1) %>% 
  filter(deviant_position != 2) %>% 
  group_by(subject, block_type) %>% 
  filter(trial_number < deviant_position) %>% 
  summarise(mean_lt = mean(trial_looking_time)) 

test_d_with_mean <- test_d %>% 
  left_join(mean_lt_d, by = c("subject", "block_type")) 

model <- lm(trial_looking_time ~ decay_rate + block_type + mean_lt , data = test_d_with_mean) 

model %>% 
  summary()

library(car)
vif(model)
```

ok maybe something slightly more fancy like residual of something???

log(looking time) ~ I((exp(1)**(-trial_number))) * item_type * trial_complexity + (trial_number * item_type * trial_complexity|subject)

```{r}
full_model <- lmerTest::lmer(log(trial_looking_time) ~ I(exp(1)**(-trial_number)) * trial_type * block_type + (1 | subject), data = d)

d$resid <- resid(full_model)

# resid at dishabituation magnitude
resid_d <- d %>% 
  filter(block_number == 1) %>% 
  filter(trial_number == deviant_position) %>% 
  filter(deviant_position != 2) %>% 
  select(subject, resid)


test_d_with_residual <- test_d %>% 
  left_join(resid_d, by = c("subject")) 



test_d_with_residual %>% 
  ggplot(aes(x = decay_rate, y = resid, color = block_type)) + 
  geom_point() + 
  theme_few() 


lm(resid ~ decay_rate + block_type  , data = test_d_with_residual)  %>% summary()


```





# Volatility measure 



```{r}
library(rugarch)

# Initialize an empty list to store GARCH volatility results
garch_volatility_list <- list()

# Group by participant and calculate each volatility measure
volatility_summary <- slope_data %>%
  group_by(subject) %>%
  summarize(
    # 1. Standard Deviation
    sd_volatility = sd(trial_looking_time, na.rm = TRUE),
    
    # 2. Mean Absolute Deviation (MAD)
    mad_volatility = mean(abs(trial_looking_time - mean(trial_looking_time, na.rm = TRUE)), na.rm = TRUE),
    
    # 3. Coefficient of Variation (CV)
    cv_volatility = sd(trial_looking_time, na.rm = TRUE) / mean(trial_looking_time, na.rm = TRUE),
    
    # 4. Range-Based Volatility
    range_volatility = max(trial_looking_time, na.rm = TRUE) - min(trial_looking_time, na.rm = TRUE),
    
    # 5. Median Absolute Deviation (MAD - Robust Version)
    mad_robust_volatility = median(abs(trial_looking_time - median(trial_looking_time, na.rm = TRUE)), na.rm = TRUE)
  )


# Calculate GARCH-based volatility for each participant and add to the volatility_summary
for (subject_id in unique(slope_data$subject)) {
  # Subset data for each participant
  participant_data <- filter(slope_data, subject == subject_id)$looking_time
  
  # Specify and fit GARCH model if there are enough data points
  if (length(participant_data) > 5) {  # Minimum data points to fit GARCH - doesn't work 
    spec <- ugarchspec(variance.model = list(model = "sGARCH"), mean.model = list(armaOrder = c(1, 1)))
    fit <- tryCatch(ugarchfit(spec = spec, data = participant_data), error = function(e) NULL)
    
    if (!is.null(fit)) {
      # Extract the average conditional volatility from the GARCH model
      avg_garch_volatility <- mean(sigma(fit), na.rm = TRUE)
    } else {
      avg_garch_volatility <- NA
    }
  } else {
    avg_garch_volatility <- NA
  }
  
  # Add the GARCH volatility to the list for each subject
  garch_volatility_list[[as.character(subject_id)]] <- avg_garch_volatility
}

# Convert GARCH volatility list to data frame and merge with other volatility measures
garch_volatility_df <- data.frame(
  subject = names(garch_volatility_list),
  garch_volatility = unlist(garch_volatility_list)
)

# Combine the GARCH results with the other volatility measures
volatility_summary <- left_join(volatility_summary, garch_volatility_df, by = "subject")

# View the final summary
print(volatility_summary)

```

```{r}
volatility_summary %>% 
  left_join(test_d %>% select(-decay_rate), by = "subject") %>% 
  pivot_longer(cols = sd_volatility: mad_robust_volatility) %>% 
  ggplot(aes(x = value, y = trial_looking_time)) + 
  geom_point(aes(color = block_type)) + 
  facet_wrap(~name, scales = "free")


volatility_summary %>% 
  left_join(test_d %>% select(-decay_rate), by = "subject") %>% 
  left_join(resid_d, by = "subject") %>% 
  pivot_longer(cols = sd_volatility: mad_robust_volatility) %>% 
  ggplot(aes(x = value, y = resid)) + 
  geom_point(aes(color = block_type)) + 
  facet_wrap(~name, scales = "free")
```

```{r}
volatility_full_d <- volatility_summary %>% 
  left_join(test_d %>% select(-decay_rate), by = "subject") %>% 
  left_join(resid_d, by = "subject") 

lm(resid ~ sd_volatility, data = volatility_full_d)  %>% summary()
lm(resid ~ mad_volatility, data = volatility_full_d)  %>% summary()
lm(resid ~ cv_volatility, data = volatility_full_d)  %>% summary()
lm(resid ~ range_volatility, data = volatility_full_d)  %>% summary()
lm(resid ~ mad_robust_volatility, data = volatility_full_d)  %>% summary()


# lol this is terrible 
all_volatility_model <- lm(resid ~ sd_volatility + mad_volatility + cv_volatility + range_volatility + mad_robust_volatility, data = volatility_full_d) 
vif(all_volatility_model)


sd_v_model <- lm(resid ~ sd_volatility, data = volatility_full_d) 
mad_v_model <- lm(resid ~ mad_volatility, data = volatility_full_d) 
cv_v_model <- lm(resid ~ cv_volatility, data = volatility_full_d)
range_v_model <- lm(resid ~ range_volatility, data = volatility_full_d) 
mad_robust_v_model <- lm(resid ~ mad_robust_volatility, data = volatility_full_d)  

models <- list(sd_v_model, mad_v_model, cv_v_model, range_v_model, mad_robust_v_model)
metrics <- sapply(models, function(model) c(Adjusted_R2 = summary(model)$adj.r.squared,
                                            AIC = AIC(model),
                                            BIC = BIC(model)))
metrics_df <- as.data.frame(t(metrics))
rownames(metrics_df) <- c("SD Volatility", "MAD Volatility", "CV Volatility", "Range Volatility", "MAD Robust Volatility")
metrics_df

```

# volatility + slope 


```{r}
v_and_slope_d <- test_d_with_residual %>% 
  left_join(volatility_full_d %>% select(-resid, -trial_looking_time, -block_type), by = "subject") 

lm(resid ~ decay_rate + sd_volatility + block_type, data = v_and_slope_d) %>% summary()
```


other features of the dataset???

- age?????
- complexity??? (social vs non-social???)



```{r}
d %>% View()
```









# Everything together 

```{r}
d %>% 
  # get rid of post dishabituation habituation trial 
  filter(trial_number < deviant_position) %>% 
  #filter(block_number == 1) %>% 
  filter(block_type == "complex_dissimilar") %>% 
  filter(deviant_position != 2) %>% 
  ungroup() %>% 
  #filter(subject %in% (d %>% distinct(subject) %>% head(10) %>% pull())) %>% 
  ggplot(aes(x = as.character(trial_number), y = trial_looking_time, group = subject, color = as.factor(deviant_position))) + 
  geom_line(alpha = .3) + 
  geom_point(alpha = .1) + 
  theme_few() + 
  facet_wrap(~block_number)
```

