---
title: "Habituation re-analysis"
output:
  html_document:
    number_sections: false
    toc: true
    toc_float: true
---



```{r}
library(tidyverse)
library(here)
library(nlme)
library(lme4)
library(rugarch)
library(ggthemes)
library(car)
```

Goal: analyzing the relationship between habituation patterns and the dishabituation magnitude across three datasets. 

- Adult dataset: https://onlinelibrary.wiley.com/doi/10.1111/tops.12631
- Preschooler dataset: https://escholarship.org/uc/item/5f38t3cd 
- Infant dataset: https://www.nature.com/articles/s41562-024-01965-x


# 1. Dataset prepping 

```{r}
adult_d <- read_csv(here("data/raw/adult_data.csv"))
preschooler_d <- read_csv(here("data/raw/preschooler_data.csv"))
infant_d <- read_csv(here("data/raw/infant_data.csv"))
```

## 1.1 Prep adult dataset 

Focusing on only the first block and blocks with more than two trials

```{r}
adult_habituation_data <- adult_d %>% 
  # fist block
  filter(block_number == 1) %>% 
  # before the dishabituation trials
  filter(trial_number < deviant_position) %>% 
  # filtering out blocks with only two trials
  filter(deviant_position != 2) %>% 
  # only keeping the relevant information 
  select(subject, trial_number, trial_looking_time) 

```




## 1.2 Prep preschooler datset 

```{r}
# figure out what blocks to retain
retained_blocks <- preschooler_d %>% 
  distinct(subject, block_number, trial_number, trial_type) %>% 
  filter(trial_type == "deviant") %>% 
  filter(trial_number > 2) %>% 
  filter(block_number == 1) %>% 
  mutate(deviant_trial_number = trial_number)


preschooler_habituation_data <- preschooler_d %>% 
  # these are participants that have the first block meeting the criteria 
  filter(subject %in% retained_blocks$subject) %>% 
  filter(block_number == 1) %>% 
  left_join(retained_blocks %>% select(-trial_number, -trial_type), by = c("subject", "block_number")) %>% 
  filter(trial_number < deviant_trial_number) %>% 
  select(subject, trial_number, trial_looking_time, age_in_months) 
```


## 1.3 Prep infant dataset 

```{r}
infant_habituation_data <- infant_d %>% 
  select(paper, subj, agedays, starts_with("train")) %>% 
  pivot_longer(cols = starts_with("train")) %>% 
  filter(!is.na(value)) %>% 
  mutate(trial_number = str_replace(name, "train", "")) %>% 
  rename(trial_looking_time = value) %>% 
  rename(subject = subj, age_in_days = agedays) %>% 
  select(paper, subject, trial_number, trial_looking_time, age_in_days) %>% 
  mutate(trial_number = as.numeric(trial_number))
```



# 2. Measure prepping 

## 2.1 Prep slope measure  

### 2.1.1 Setting up helper function 

```{r}

# running the information 
get_individual_slope <- function(data){
  
  mixed_exp_model <- nlme(
  trial_looking_time ~ a * exp(b * trial_number),  # Simplified exponential decay model
  data = data,
  fixed = a + b ~ 1,                # Fixed effects for a and b
  random = b ~ 1 | subject,         # Random effect for b (decay rate) by subject
  start = c(a = max(data$trial_looking_time), 
            b = -0.05)  # Starting values
  )
  
  # Extract random effects (individual decay rates)
  random_effects <- ranef(mixed_exp_model)$b

  # Combine fixed effect with random effect for individual slopes
  fixed_decay_rate <- fixef(mixed_exp_model)["b"]
  individual_decay_rates <- fixed_decay_rate + random_effects
  
  # Create a data frame to store the slopes for each participant
  slopes <- data.frame(
    subject = rownames(ranef(mixed_exp_model)),
    decay_rate = individual_decay_rates
  )
  
  
  mixed_exp_model <- nlme(
  trial_looking_time ~ a * exp(b * trial_number),  # Simplified exponential decay model
  data = data,
  fixed = a + b ~ 1,                # Fixed effects for a and b
  random = a + b ~ 1 | subject,         # Random effect for b (decay rate) by subject
  start = c(a = max(data$trial_looking_time), 
            b = -0.05)  # Starting values
  )
  
  # Extract random effects (individual decay rates)
  random_effects <- ranef(mixed_exp_model)$b
  fixed_decay_rate <- fixef(mixed_exp_model)["b"]
  individual_decay_rates <- fixed_decay_rate + random_effects

  # Create a data frame to store the slopes for each participant
  slopes$decay_rates_random <-individual_decay_rates
  
  return(slopes)
}



```

### 2.1.2 Demo how the curve chnages shape 

```{r}

# Define a function for the exponential decay
decay_model <- function(trial_number, a, b) {
  a * exp(b * trial_number)
}

# Create a data frame with different combinations of a and b
trial_data <- expand.grid(
  trial_number = seq(0, 10, by = 0.1),
  a = c(10, 20, 30),  # Different initial values
  b = c(-0.1, -0.2, -0.3, 0.1)  # Different decay rates
)

# Calculate the looking time based on the decay model
trial_data <- trial_data %>%
  mutate(trial_looking_time = decay_model(trial_number, a, b))

# Plot the curves
ggplot(trial_data, aes(x = trial_number, y = trial_looking_time, color = factor(a), linetype = factor(b))) +
  geom_line(size = 1) +
  labs(
    title = "Effect of Parameters 'a' and 'b' on Exponential Decay Curve",
    x = "Trial Number",
    y = "Trial Looking Time",
    color = "Initial Value (a)",
    linetype = "Decay Rate (b)"
  ) +
  theme_minimal()

```

### 2.1.3 Fitting the models to all the datasets

```{r}
adult_slope_df <- get_individual_slope(adult_habituation_data)
preschooler_slope_df <- get_individual_slope(preschooler_habituation_data)
infant_slope_df <- get_individual_slope(infant_habituation_data)
```


## 2.2 Prep volatility measure 

### 2.2.1 Setting up helper function 

```{r}

get_volatility_measures <- function(data){
  
  # Initialize an empty list to store GARCH volatility results
garch_volatility_list <- list()

# Group by participant and calculate each volatility measure
volatility_summary <- data %>%
  group_by(subject) %>%
  summarize(
    # 1. Standard Deviation
    sd_volatility = sd(trial_looking_time, na.rm = TRUE),
    
    # 2. Mean Absolute Deviation (MAD)
    mad_volatility = mean(abs(trial_looking_time - mean(trial_looking_time, na.rm = TRUE)), na.rm = TRUE),
    
    # 3. Coefficient of Variation (CV)
    cv_volatility = sd(trial_looking_time, na.rm = TRUE) / mean(trial_looking_time, na.rm = TRUE),
    
    # 4. Range-Based Volatility
    range_volatility = max(trial_looking_time, na.rm = TRUE) - min(trial_looking_time, na.rm = TRUE),
    
    # 5. Median Absolute Deviation (MAD - Robust Version)
    mad_robust_volatility = median(abs(trial_looking_time - median(trial_looking_time, na.rm = TRUE)), na.rm = TRUE)
  )


# Calculate GARCH-based volatility for each participant and add to the volatility_summary
for (subject_id in unique(data$subject)) {
  # Subset data for each participant
  participant_data <- filter(data, subject == subject_id)$trial_looking_time
  
  # Specify and fit GARCH model if there are enough data points
  if (length(participant_data) > 5) {  # Minimum data points to fit GARCH - doesn't work 
    spec <- ugarchspec(variance.model = list(model = "sGARCH"), mean.model = list(armaOrder = c(1, 1)))
    fit <- tryCatch(ugarchfit(spec = spec, data = participant_data), error = function(e) NULL)
    
    if (!is.null(fit)) {
      # Extract the average conditional volatility from the GARCH model
      avg_garch_volatility <- mean(sigma(fit), na.rm = TRUE)
    } else {
      avg_garch_volatility <- NA
    }
  } else {
    avg_garch_volatility <- NA
  }
  
  # Add the GARCH volatility to the list for each subject
  garch_volatility_list[[as.character(subject_id)]] <- avg_garch_volatility
}

# Convert GARCH volatility list to data frame and merge with other volatility measures
garch_volatility_df <- data.frame(
  subject = names(garch_volatility_list),
  garch_volatility = unlist(garch_volatility_list)
)

# Combine the GARCH results with the other volatility measures
volatility_summary <- left_join(volatility_summary, garch_volatility_df, by = "subject")
  
return(volatility_summary)
}



```


### 2.2.2  Fit with each dataset

```{r}
adult_volatility_measure <- get_volatility_measures(adult_habituation_data)
preschooler_volatility_measure <-  get_volatility_measures(preschooler_habituation_data)
infant_volatility_measure <-  get_volatility_measures(infant_habituation_data)
```

## 2.3 Put together all the measures 

```{r}
adult_measures <- adult_volatility_measure %>% 
  left_join(adult_slope_df, by = c("subject"))

preschooler_measures <- preschooler_volatility_measure %>% 
  left_join(preschooler_slope_df, by = c("subject"))

infant_measures <- infant_volatility_measure %>% 
  left_join(infant_slope_df, by = c("subject"))
```


# 3. Dishabituation and measure dataset prepping 

## 3.1 Adult 


### 3.1.1 get residual 

```{r}
adult_full_raw <- adult_d %>% 
  filter(block_number == 1) 

adult_full_raw$trial_type_numeric <- as.numeric(as.factor(adult_full_raw$trial_type))

habituation_model <- nlme(
  trial_looking_time ~ a * exp(b * trial_number * trial_type_numeric) ,  # Simplified exponential decay model
  data = adult_full_raw,
  fixed = a + b ~ 1,                # Fixed effects for a and b
  random = a + b ~ 1 | subject,         # Random effect for b (decay rate) by subject
  start = c(a = max(adult_full_raw$trial_looking_time), 
            b = -0.05)
)

adult_full_raw$residualized_looking_time <- resid(habituation_model)
```




#### 3.1.1.1 model checking

```{r}
# Check residual summary statistics
summary(adult_full_raw$residualized_looking_time)

# Histogram of residuals
hist(adult_full_raw$residualized_looking_time, breaks = 30,
     main = "Histogram of Residualized Looking Time", xlab = "Residuals")


plot(adult_full_raw$trial_looking_time, adult_full_raw$residualized_looking_time,
     xlab = "Raw Looking Time", ylab = "Residualized Looking Time",
     main = "Residuals vs Raw Looking Time")+
abline(h = 0, col = "red", lty = 2)


cor(adult_full_raw$trial_looking_time, adult_full_raw$residualized_looking_time, use = "complete.obs")

# Add predicted values to the dataset
adult_full_raw$fitted_values <- predict(habituation_model)

# Visualize individual fits
ggplot(adult_full_raw %>% filter(subject %in% sample(adult_full_raw %>% distinct(subject) %>% pull(), 20)), aes(x = trial_number, y = trial_looking_time, color = trial_type)) +
  geom_point(alpha = 0.5) +
  geom_line(aes(y = fitted_values), size = 1) +
  facet_wrap(~ subject, scales = "free_y") +
  labs(title = "Observed vs Fitted Values for Each Participant",
       x = "Trial Number", y = "Looking Time") +
  theme_minimal()
```





### 3.1.2 get difference 

```{r}
# resid at dishabituation magnitude
adult_dishab_raw <- adult_full_raw %>% 
  filter(block_number == 1) %>% 
  filter(trial_number == deviant_position) %>% 
  filter(deviant_position != 2) 

adult_dishab_d <- adult_dishab_raw %>% 
  select(subject, trial_looking_time) %>% 
  rename(dishab_raw = trial_looking_time) %>% 
  left_join(
    adult_full_raw %>%  filter(trial_number < deviant_position) %>% group_by(subject) %>% filter(trial_number == max(trial_number)),
    by = c("subject")
  ) %>% 
  mutate(dishab_diff = log(dishab_raw) - log(trial_looking_time)) %>% 
  mutate(dishab_trial_num = trial_number + 1) %>% 
  select(subject, dishab_diff, dishab_trial_num)

```


```{r}
adult_dishab_d

adult_habituation_data %>% 
  group_by(subject) %>% 
summarise(avg_lt = mean(trial_looking_time)) %>% 
  left_join(adult_dishab_d, by = c("subject")) %>% 
  ggplot(aes(x = avg_lt, y = dishab_diff)) + 
  geom_point() + 
  geom_smooth(method = "lm")
```


### 3.1.3 link age 

get the mapping between prolific ID and subject ID to extract the age information 

```{r}
raw_adult <- read_csv(here("data/raw/adult_raw_with_prolific.csv"))
prolific_d <- read_csv(here("data/raw/adult_prolific.csv"))

subject_prolific_link <- raw_adult %>% 
  filter(grepl("p_id", responses)) %>% 
  mutate(prolific_id = str_extract(responses, '(?<="p_id":")[^"]+')) %>% 
  select(subject, prolific_id)

link_df <- prolific_d %>% 
  rename(prolific_id = participant_id) %>% 
  select(prolific_id, age) %>% 
  left_join(subject_prolific_link, by = c("prolific_id"))

adult_dishab_d <- adult_dishab_d %>% 
  left_join(link_df %>% select(-prolific_id), by = c("subject")) 
```


### 3.1.3 combine with the measures 

```{r}
adult_measures_with_dishab <- adult_dishab_d %>% 
  left_join(adult_measures, by = c("subject"))
```



## 3.2 Preschooler 

### 3.2.1. Residual 

```{r}
preschooler_full_raw <- preschooler_d %>% 
  filter(block_number == 1) 

preschooler_full_raw$trial_type_numeric <- as.numeric(as.factor(preschooler_full_raw$trial_type))

habituation_model <- nlme(
  trial_looking_time ~ a * exp(b * trial_number * trial_type_numeric) ,  # Simplified exponential decay model
  data =  preschooler_full_raw,
  fixed = a + b ~ 1,                # Fixed effects for a and b
  random = a + b ~ 1 | subject,         # Random effect for b (decay rate) by subject
  start = c(a = max(preschooler_full_raw$trial_looking_time), 
            b = -0.05)
)


preschooler_full_raw$residualized_looking_time <- resid(habituation_model)
```


#### 3.2.1.1 Model checking

```{r}
# Check residual summary statistics
summary(preschooler_full_raw$residualized_looking_time)

# Histogram of residuals
hist(preschooler_full_raw$residualized_looking_time, breaks = 30,
     main = "Histogram of Residualized Looking Time", xlab = "Residuals")


plot(preschooler_full_raw$trial_looking_time, preschooler_full_raw$residualized_looking_time,
     xlab = "Raw Looking Time", ylab = "Residualized Looking Time",
     main = "Residuals vs Raw Looking Time")+
abline(h = 0, col = "red", lty = 2)


cor(preschooler_full_raw$trial_looking_time, preschooler_full_raw$residualized_looking_time, use = "complete.obs")

# Add predicted values to the dataset
preschooler_full_raw$fitted_values <- predict(habituation_model)

# Visualize individual fits
ggplot(preschooler_full_raw %>% filter(subject %in% sample(preschooler_full_raw %>% distinct(subject) %>% pull(), 20)), aes(x = trial_number, y = trial_looking_time, color = trial_type)) +
  geom_point(alpha = 0.5) +
  geom_line(aes(y = fitted_values), size = 1) +
  facet_wrap(~ subject, scales = "free_y") +
  labs(title = "Observed vs Fitted Values for Each Participant",
       x = "Trial Number", y = "Looking Time") +
  theme_minimal()
```




### 3.2.2 Put together 


```{r}
#preschooler_full_model <- lmerTest::lmer(log(trial_looking_time) ~ I(exp(1)**(-trial_number)) * trial_type  + (1 | subject), data = preschooler_d)
#preschooler_d$resid <- resid(preschooler_full_model)

# resid at dishabituation magnitude
preschooler_dishab_d <- preschooler_full_raw %>% 
  filter(subject %in% retained_blocks$subject) %>% 
  filter(block_number == 1) %>% 
  left_join(retained_blocks %>% select(-trial_number, -trial_type), by = c("subject", "block_number")) %>% 
  filter(trial_number == deviant_trial_number) %>% 
  rename(dishab_raw = residualized_looking_time) %>% 
  left_join(
    preschooler_full_raw %>% group_by(subject) %>%  filter(subject %in% retained_blocks$subject) %>% 
  filter(block_number == 1) %>% 
  left_join(retained_blocks %>% select(-trial_number, -trial_type), by = c("subject", "block_number")) %>% 
  filter(trial_number < deviant_trial_number) %>%
  group_by(subject) %>% 
  filter(trial_number == max(trial_number)) %>% 
    ungroup() %>% 
  select(subject,  residualized_looking_time), 
    by = c("subject")
  ) %>% 
  mutate(dishab_diff = dishab_raw - trial_looking_time) %>% 
  mutate(dishab_trial_num = trial_number + 1) %>% 
  select(subject, dishab_diff, dishab_trial_num, age_in_months)
  
preschooler_measures_with_dishab <- preschooler_dishab_d %>% 
  left_join(preschooler_measures, by = c("subject"))
```


## 3.3 Infants 

### 3.1.1 Re-wrangle some datasets 


```{r}
# try to construct the dataset to mimic the other ones, using only first of the unexpected as the "dishabituating trials"
# figure out the trial number for the extra trial
last_trial_number <- infant_d %>% 
  select(paper, subj, agedays, starts_with("train"), "expected1") %>% 
  rename(trainEXTRA = expected1) %>% 
  pivot_longer(cols = starts_with("train")) %>% 
  filter(!is.na(value)) %>% 
  mutate(trial_number = as.numeric(str_replace(name, "train", ""))) %>% 
 group_by(paper, subj) %>% 
  summarise(max_trial = max(trial_number, na.rm = TRUE)) %>% 
  mutate(last_trial = max_trial + 1) %>% 
  select(paper, subj, last_trial) %>% 
  ungroup() %>% 
  filter(!last_trial == -Inf)

```


### 3.1.2 Residuals 

```{r}
#preschooler_full_model <- lmerTest::lmer(log(trial_looking_time) ~ I(exp(1)**(-trial_number)) * trial_type  + (1 | subject), data = preschooler_d)

infant_model_d <- infant_d %>% 
  select(paper, subj, agedays, starts_with("train"), "expected1") %>% 
  rename(trainEXTRA = expected1) %>% 
  pivot_longer(cols = starts_with("train")) %>% 
  filter(!is.na(value)) %>% 
  mutate(trial_number = as.numeric(str_replace(name, "train", ""))) %>% 
  left_join(last_trial_number, by = c("paper", "subj")) %>% 
  mutate(trial_type = case_when(
    is.na(trial_number) ~ "deviant", 
    TRUE ~ "background"
  )) %>% 
  mutate(trial_number = case_when(
    is.na(trial_number) ~ last_trial, 
    TRUE ~ trial_number
  )) %>% 
  filter(!is.na(trial_number)) %>% 
  select(paper, subj, trial_number, trial_type, value, agedays) %>% 
  rename(trial_looking_time = value, 
         subject = subj)

infant_model_d$trial_type_numeric <- as.numeric(as.factor(infant_model_d$trial_type))

habituation_model <- nlme(
  trial_looking_time ~ a * exp(b * trial_number * trial_type_numeric) ,  # Simplified exponential decay model
  data =  infant_model_d,
  fixed = a + b ~ 1,                # Fixed effects for a and b
  random = a + b ~ 1 | subject,         # Random effect for b (decay rate) by subject
  start = c(a = max(infant_model_d$trial_looking_time), 
            b = -0.05)
)


infant_model_d$residualized_looking_time <- resid(habituation_model)
```
#### 3.1.2.1 model checking

```{r}
# Check residual summary statistics
summary(infant_model_d$residualized_looking_time)

# Histogram of residuals
hist(infant_model_d$residualized_looking_time, breaks = 30,
     main = "Histogram of Residualized Looking Time", xlab = "Residuals")


plot(infant_model_d$trial_looking_time, infant_model_d$residualized_looking_time,
     xlab = "Raw Looking Time", ylab = "Residualized Looking Time",
     main = "Residuals vs Raw Looking Time")+
abline(h = 0, col = "red", lty = 2)


cor(infant_model_d$trial_looking_time, infant_model_d$residualized_looking_time, use = "complete.obs")

# Add predicted values to the dataset
infant_model_d$fitted_values <- predict(habituation_model)

# Visualize individual fits
ggplot(infant_model_d %>% filter(subject %in% sample(infant_model_d %>% distinct(subject) %>% pull(), 20)), aes(x = trial_number, y = trial_looking_time, color = trial_type)) +
  geom_point(alpha = 0.5) +
  geom_line(aes(y = fitted_values), size = 1) +
  facet_wrap(~ subject, scales = "free_y") +
  labs(title = "Observed vs Fitted Values for Each Participant",
       x = "Trial Number", y = "Looking Time") +
  theme_minimal()
```



### 3.1.3  Put things together 

```{r}
duplicate_subjects <- infant_model_d %>% group_by(paper, subject) %>% filter(trial_type != "deviant") %>%  filter(trial_number == max(trial_number)) %>% select(paper, subject, trial_number, residualized_looking_time, agedays) %>% group_by(paper, subject) %>% count() %>% filter(n > 1)
```


```{r}
infant_dishab_d <- infant_model_d %>% 
  filter(trial_type == "deviant") %>% 
  filter(!(paper %in% duplicate_subjects$paper & subject %in% duplicate_subjects$subject)) %>% 
  select(paper, subject, residualized_looking_time) %>% 
  rename(dishab_raw = residualized_looking_time) %>% 
  left_join(
    infant_model_d %>% group_by(paper, subject) %>% filter(trial_type != "deviant") %>%  filter(trial_number == max(trial_number)) %>% select(paper, subject, trial_number, residualized_looking_time, agedays) %>% filter(!(paper %in% duplicate_subjects$paper & subject %in% duplicate_subjects$subject)), 
    by = c("paper", "subject")
  ) %>% 
  mutate(dishab_diff = dishab_raw - residualized_looking_time) %>% 
   mutate(dishab_trial_num = trial_number + 1) %>% 
  select(paper, subject, dishab_diff, dishab_trial_num, agedays)

```


```{r}
infant_measure_with_dishab <- infant_dishab_d %>% 
  select(paper, subject, dishab_trial_num, dishab_diff, agedays) %>% 
  left_join(infant_measures, by = c("subject"))

```


# 4. Visualization 

## 4.1. Raw habituation data Visualizatino 

### 4.1.1 Adults

```{r}
adult_habituation_data %>% 
  ggplot(aes(x = trial_number, y = trial_looking_time, group = subject)) + 
  geom_point(alpha = .1) + 
  geom_line(alpha = .1) 
 

adult_habituation_data %>% 
  ggplot(aes(x = trial_number, y = trial_looking_time, group = subject)) + 
  geom_point(alpha = .1) + 
  geom_line(alpha = .1) +
  scale_y_log10()

adult_habituation_data %>% 
  ggplot(aes(x = trial_number, y = trial_looking_time)) + 
  geom_point(alpha = .1) + 
  stat_summary(fun.data = "mean_cl_boot", geom = "pointrange", color = "red") 
 

adult_habituation_data %>% 
  ggplot(aes(x = trial_number, y = trial_looking_time)) + 
  geom_point(alpha = .1) + 
  stat_summary(fun.data = "mean_cl_boot", geom = "pointrange", color = "red") +
  scale_y_log10()
```



### 4.1.2 Preschoolers

```{r}
preschooler_habituation_data %>% 
  ggplot(aes(x = trial_number, y = trial_looking_time, group = subject)) + 
  geom_point(alpha = .1) + 
  geom_line(alpha = .1) 
 

preschooler_habituation_data %>% 
  ggplot(aes(x = trial_number, y = trial_looking_time, group = subject)) + 
  geom_point(alpha = .1) + 
  geom_line(alpha = .1) +
  scale_y_log10()

preschooler_habituation_data %>% 
  ggplot(aes(x = trial_number, y = trial_looking_time)) + 
  geom_point(alpha = .1) + 
  stat_summary(fun.data = "mean_cl_boot", geom = "pointrange", color = "red") 
 

preschooler_habituation_data %>% 
  ggplot(aes(x = trial_number, y = trial_looking_time)) + 
  geom_point(alpha = .1) + 
  stat_summary(fun.data = "mean_cl_boot", geom = "pointrange", color = "red") +
  scale_y_log10()
```



### 4.1.3 Infants

```{r}
infant_habituation_data %>% 
  mutate(individual_infant = paste0(paper, subject)) %>% 
  ggplot(aes(x = trial_number, y = trial_looking_time, group = individual_infant)) + 
  geom_point(alpha = .1) + 
  geom_line(alpha = .1) 
 

infant_habituation_data %>% 
  mutate(individual_infant = paste0(paper, subject)) %>% 
  ggplot(aes(x = trial_number, y = trial_looking_time, group = individual_infant))  + 
  geom_point(alpha = .1) + 
  geom_line(alpha = .1) +
  scale_y_log10()

infant_habituation_data %>% 
  ggplot(aes(x = trial_number, y = trial_looking_time)) + 
  geom_point(alpha = .1) + 
  stat_summary(fun.data = "mean_cl_boot", geom = "pointrange", color = "red") 
 

infant_habituation_data %>% 
  ggplot(aes(x = trial_number, y = trial_looking_time)) + 
  geom_point(alpha = .1) + 
  stat_summary(fun.data = "mean_cl_boot", geom = "pointrange", color = "red") +
  scale_y_log10()
```



# 4.2 Raw dishabituation data visualization 

```{r}
adult_measures_with_dishab %>% 
  ggplot(aes(x = dishab_diff)) + 
  geom_density(alpha = .1, aes(fill = as.factor(dishab_trial_num)))
```

```{r}
preschooler_measures_with_dishab %>% 
  ggplot(aes(x = dishab_diff)) + 
  geom_density(alpha = .1, aes(fill = as.factor(dishab_trial_num)))
```

```{r}
infant_measure_with_dishab %>% 
  ggplot(aes(x = dishab_diff)) + 
  geom_density()
```




# 4.3 Measure and data predictions 


```{r}
full_d <- bind_rows(
  infant_measure_with_dishab %>% mutate(group = "infants"), 
  preschooler_measures_with_dishab %>% mutate(group = "preschoolers"), 
  adult_measures_with_dishab %>% mutate(group = "adults")
)
```

### 4.3.1 Slope measures vs resid across age groups 

```{r}
full_d %>% 
  ggplot(aes(x = decay_rate, y = dishab_diff)) + 
  geom_point(alpha = .2) + 
  facet_wrap(~group, scales = "free") + 
  geom_smooth(method = "lm") 

full_d %>% 
  select(decay_rate, decay_rates_random, dishab_diff, group) %>% 
  pivot_longer(cols = c("decay_rate", "decay_rates_random"), names_to = "decay_rate_type", values_to = "val") %>% 
  ggplot(aes(x = val, y = dishab_diff)) + 
  geom_point(alpha = .2) + 
  facet_wrap(decay_rate_type~group, scales = "free") + 
  geom_smooth(method = "lm") 


# log would remove the negative ones 
full_d %>% 
  ggplot(aes(x = decay_rate, y = log(dishab_diff))) + 
  geom_point(alpha = .2) + 
  facet_wrap(~group, scales = "free") + 
  geom_smooth(method = "lm") 
```

## 4.3.2 Volatility measurs across age groups 



```{r fig.width=8, fig.height=8}
full_d %>% 
  pivot_longer(cols = ends_with("volatility"), names_to = "volatility_type", values_to =  "volatility_measure") %>% 
  ggplot(aes(x = volatility_measure, y = dishab_diff)) + 
  geom_point(alpha = .2) + 
  facet_wrap(volatility_type~group, scales = "free") + 
  geom_smooth(method = "lm") 
```


# 5. Models 

## 5.1 Adults 

### 5.1.2 Decay Rate only 

```{r}
lm(dishab_diff ~ decay_rate, data = adult_measures_with_dishab) %>% summary()
```

```{r}
lm(dishab_diff ~ decay_rates_random , data = adult_measures_with_dishab) %>% summary()

```

### 5.1.3 Decay Rate + Age 

```{r}
lm(dishab_diff ~ decay_rate + age, data = adult_measures_with_dishab) %>% summary()
```

```{r}
lm(dishab_diff ~ decay_rates_random + age, data = adult_measures_with_dishab) %>% summary()

```


### 5.1.3 Decay Rate + Age + volatility measure

```{r}
lm(dishab_diff ~ decay_rate + age + sd_volatility, data = adult_measures_with_dishab) %>% summary()
```

```{r}
lm(dishab_diff ~ decay_rates_random + age + sd_volatility, data = adult_measures_with_dishab) %>% summary()
```

## 5.2 Preschoolers 

### 5.1.2 Decay Rate only 

```{r}
lm(dishab_diff ~ decay_rate , data = preschooler_measures_with_dishab) %>% summary()
```

```{r}
lm(dishab_diff ~ decay_rates_random , data = preschooler_measures_with_dishab) %>% summary()
```

### 5.1.3 Decay Rate + Age 

```{r}
lm(dishab_diff ~ decay_rate + age_in_months, data = preschooler_measures_with_dishab) %>% summary()
```

```{r}
lm(dishab_diff ~ decay_rates_random + age_in_months, data = preschooler_measures_with_dishab) %>% summary()

```

### 5.1.3 Decay Rate + Age + volatility measure

```{r}
lm(dishab_diff ~ decay_rate + age_in_months + sd_volatility, data = preschooler_measures_with_dishab) %>% summary()
```


```{r}
vif(lm(dishab_diff ~ decay_rate + age_in_months + sd_volatility, data = preschooler_measures_with_dishab))
```

```{r}
lm(dishab_diff ~ decay_rates_random + age_in_months + sd_volatility, data = preschooler_measures_with_dishab) %>% summary()
```


## 6.1 Infants 

### 6.1.1 Decay Rate only 

```{r}
lm(dishab_diff ~ decay_rate , data = infant_measure_with_dishab) %>% summary()
```

```{r}
lm(dishab_diff ~ decay_rates_random , data = infant_measure_with_dishab) %>% summary()
```

### 5.1.3 Decay Rate + Age 

```{r}
lm(dishab_diff ~ decay_rate + agedays, data = infant_measure_with_dishab) %>% summary()
```

```{r}
lm(dishab_diff ~ decay_rates_random + agedays, data = infant_measure_with_dishab) %>% summary()
```

### 5.1.3 Decay Rate + Age + volatility measure

```{r}
lm(dishab_diff ~ decay_rate + agedays + sd_volatility, data = infant_measure_with_dishab) %>% summary()
```

```{r}
lm(dishab_diff ~ decay_rates_random + agedays + sd_volatility, data = infant_measure_with_dishab) %>% summary()

```


# 6. Summarizing Model results

## 6.1 Combine Decay Rate Only models

```{r}

bind_rows(
  lm(dishab_diff ~ decay_rate, data = adult_measures_with_dishab) %>% broom::tidy(conf.int = TRUE, conf.level = 0.95) %>% mutate(group = "adults"),
  lm(dishab_diff ~ decay_rate, data = preschooler_measures_with_dishab) %>% broom::tidy(conf.int = TRUE, conf.level = 0.95) %>% mutate(group = "preschoolers"),
  lm(dishab_diff ~ decay_rate, data = infant_measure_with_dishab) %>% broom::tidy(conf.int = TRUE, conf.level = 0.95) %>% mutate(group = "infants")
) %>% 
  filter(term == "decay_rate") %>% 
  ggplot(aes(x = factor(group, levels = c("infants", "preschoolers", "adults")), y = estimate, ymin = conf.low, ymax = conf.high)) + 
  geom_pointrange() + 
  geom_hline(yintercept = 0, color = "gray", linetype = "dashed") + 
  coord_flip() + 
  theme_few() + 
  xlab("")


```


```{r}
bind_rows(
  lm(dishab_diff ~ decay_rates_random, data = adult_measures_with_dishab) %>% broom::tidy(conf.int = TRUE, conf.level = 0.95) %>% mutate(group = "adults"),
  lm(dishab_diff ~ decay_rates_random, data = preschooler_measures_with_dishab) %>% broom::tidy(conf.int = TRUE, conf.level = 0.95) %>% mutate(group = "preschoolers"),
  lm(dishab_diff ~ decay_rates_random, data = infant_measure_with_dishab) %>% broom::tidy(conf.int = TRUE, conf.level = 0.95) %>% mutate(group = "infants")
) %>% 
  filter(term == "decay_rates_random") %>% 
  ggplot(aes(x = factor(group, levels = c("infants", "preschoolers", "adults")), y = estimate, ymin = conf.low, ymax = conf.high)) + 
  geom_pointrange() + 
  geom_hline(yintercept = 0, color = "gray", linetype = "dashed") + 
  coord_flip() + 
  theme_few() + 
  xlab("")
```

## 6.2 Combine Decay Rate + Age Models

```{r}
bind_rows(
   lm(dishab_diff ~ decay_rate + age, data = adult_measures_with_dishab) %>% broom::tidy(conf.int = TRUE, conf.level = 0.95) %>% mutate(group = "adults"),
  lm(dishab_diff ~ decay_rate + age_in_months, data = preschooler_measures_with_dishab) %>% broom::tidy(conf.int = TRUE, conf.level = 0.95) %>% mutate(group = "preschoolers"),
  lm(dishab_diff ~ decay_rate + agedays, data = infant_measure_with_dishab) %>% broom::tidy(conf.int = TRUE, conf.level = 0.95) %>% mutate(group = "infants")
) %>% 
  filter(term != "(Intercept)") %>% 
  mutate(clean_term = case_when(
    term == "age_in_months" ~ "age",
    term == "agedays" ~ "age", 
    TRUE ~ term
)) %>% 
  ggplot(aes(x = factor(group, levels = c("infants", "preschoolers", "adults")), y = estimate, ymin = conf.low, ymax = conf.high)) + 
  geom_pointrange() + 
  geom_hline(yintercept = 0, color = "gray", linetype = "dashed") + 
  coord_flip() + 
  theme_few() + 
  xlab("") + 
  facet_wrap(~clean_term)
```


```{r}
bind_rows(
   lm(dishab_diff ~ decay_rates_random + age, data = adult_measures_with_dishab) %>% broom::tidy(conf.int = TRUE, conf.level = 0.95) %>% mutate(group = "adults"),
  lm(dishab_diff ~ decay_rates_random + age_in_months, data = preschooler_measures_with_dishab) %>% broom::tidy(conf.int = TRUE, conf.level = 0.95) %>% mutate(group = "preschoolers"),
  lm(dishab_diff ~ decay_rates_random + agedays, data = infant_measure_with_dishab) %>% broom::tidy(conf.int = TRUE, conf.level = 0.95) %>% mutate(group = "infants")
) %>% 
  filter(term != "(Intercept)") %>% 
  mutate(clean_term = case_when(
    term == "age_in_months" ~ "age",
    term == "agedays" ~ "age", 
    TRUE ~ term
)) %>% 
  ggplot(aes(x = factor(group, levels = c("infants", "preschoolers", "adults")), y = estimate, ymin = conf.low, ymax = conf.high)) + 
  geom_pointrange() + 
  geom_hline(yintercept = 0, color = "gray", linetype = "dashed") + 
  coord_flip() + 
  theme_few() + 
  xlab("") + 
  facet_wrap(~clean_term)
```

## 6.3 Combine Decay Rate + Age + Volatility Models

```{r}
bind_rows(
   lm(dishab_diff ~ decay_rate + age + sd_volatility, data = adult_measures_with_dishab) %>% broom::tidy(conf.int = TRUE, conf.level = 0.95) %>% mutate(group = "adults"),
  lm(dishab_diff ~ decay_rate + age_in_months + sd_volatility, data = preschooler_measures_with_dishab) %>% broom::tidy(conf.int = TRUE, conf.level = 0.95) %>% mutate(group = "preschoolers"),
  lm(dishab_diff ~ decay_rate + agedays + sd_volatility, data = infant_measure_with_dishab) %>% broom::tidy(conf.int = TRUE, conf.level = 0.95) %>% mutate(group = "infants")
) %>% 
  filter(term != "(Intercept)") %>% 
  mutate(clean_term = case_when(
    term == "age_in_months" ~ "age",
    term == "agedays" ~ "age", 
    TRUE ~ term
)) %>% 
  ggplot(aes(x = factor(group, levels = c("infants", "preschoolers", "adults")), y = estimate, ymin = conf.low, ymax = conf.high)) + 
  geom_pointrange() + 
  geom_hline(yintercept = 0, color = "gray", linetype = "dashed") + 
  coord_flip() + 
  theme_few() + 
  xlab("") + 
  facet_wrap(~clean_term, scales = "free")
```

```{r}
bind_rows(
   lm(dishab_diff ~ decay_rates_random + age + sd_volatility, data = adult_measures_with_dishab) %>% broom::tidy(conf.int = TRUE, conf.level = 0.95) %>% mutate(group = "adults"),
  lm(dishab_diff ~ decay_rates_random + age_in_months + sd_volatility, data = preschooler_measures_with_dishab) %>% broom::tidy(conf.int = TRUE, conf.level = 0.95) %>% mutate(group = "preschoolers"),
  lm(dishab_diff ~ decay_rates_random + agedays + sd_volatility, data = infant_measure_with_dishab) %>% broom::tidy(conf.int = TRUE, conf.level = 0.95) %>% mutate(group = "infants")
) %>% 
  filter(term != "(Intercept)") %>% 
  mutate(clean_term = case_when(
    term == "age_in_months" ~ "age",
    term == "agedays" ~ "age", 
    TRUE ~ term
)) %>% 
  ggplot(aes(x = factor(group, levels = c("infants", "preschoolers", "adults")), y = estimate, ymin = conf.low, ymax = conf.high)) + 
  geom_pointrange() + 
  geom_hline(yintercept = 0, color = "gray", linetype = "dashed") + 
  coord_flip() + 
  theme_few() + 
  xlab("") + 
  facet_wrap(~clean_term, scales = "free")
```


